```{r memo}

library("rethinking")
# Causal effects
library(dagitty)
kdag<-dagitty("dag {
  
  S-> K <- T 
  T -> S
  
  T -> SL -> t


}")
adjustmentSets(kdag, exposure="SL", outcome = "K")



library(dagitty)
kdag<-dagitty("dag {
  G->Y<-X 
  G->X 
  Z->Y



}")
adjustmentSets(kdag, exposure="X", outcome = "Y")



T<-1:25
SL<-NULL
for (i in T)SL[i]<- rnorm(1, - (T[i] - 50)^2 * 2/ 800+ 3, 0.6)

T<- round(runif(200, 1,25))
SL<-NULL
for (i in 1:length(T)) SL[i]<- rnorm(1, - (1/ 40 )*(T[i] - 12.5)^2 + 2, 0.5)
plot(T, SL)

dS<-rbeta(length(T), 1.2,1)
S<-NULL
for(i in 1:length(T_K))S[i] <- rnorm(1, 0.01*T[i]-2.2*dS[i]+5, 0.4)/7.5
plot(density(S))
plot(dS, S)
curve((-2.2 * x + 0.01*mean(T)+ 5)/7.5, add=TRUE)

dK<-rbeta(length(T), 1.4,1)
K<-NULL
for(i in 1:length(T))K[i] <- rnorm(1, 1.5*SL[i] + 0.01* T[i] + 3.2* S[i]- 2*dK[i]+7, 0.1)/17
plot(density(K))

# check
plot(dK, K)
curve((-2 * x + 1.5*mean(SL) + 0.01* mean(T) + 3.2* mean(S)+ 7)/17, add=TRUE)
plot(SL, K)
curve((-2 * mean(dS) + 1.5*x + 0.01* mean(T) + 3.2* mean(S)+ 7)/17, add=TRUE)
plot(S, K)
curve((-2 * mean(dS) + 1.5*mean(SL) + 0.01* mean(T) + 3.2* x+ 7)/17, add=TRUE)

d<-list(T=T, SL=SL, dS=dS, S=S, K=K, dK=dK )
d


plot(K, exp()/(1+exp()))
m1<- ulam(
    alist (
        K~dnorm(mu, sigma), 
        mu <- a[T] + b_SL[T] * SL, 
        a[T]~dnorm(0,1), 
        b_SL[T]~dexp(1), 
        sigma~dexp(1)

    ), dat = list(K=d$K, SL=d$SL, T=d$T, S=d$S), chains=4
)
precis(m1, depth=2)
dashboard(m1)
# Plots

plot(SL, K)
curve(0.45 + 0.09*x, add=TRUE) # no pooling 

# partial poling with one parameter
points(SL[d$T == 2], K[d$T == 2], col =2, lwd = 4)
curve(0.40 + 0.05*x, col = 2, add=TRUE, lwd=3)
points(SL[d$T == 24], K[d$T == 24], col =3, lwd = 4)
curve(0.43 + 0.07*x, col = 3, add=TRUE, lwd=3)

# Posterior Prediction
plot(SL, K)
post1<-sim(m1, data = list(SL = seq(-3, 3, len=20), T = rep(2, 20))) # T = 2 
means1<-apply(post1, 2, mean)
PIs1<-apply(post1, 2, PI)
lines(seq(-3, 3, len=20), means1, col = 2)
lines(seq(-3, 3, len=20), PIs1[1,], col = 2)
lines(seq(-3, 3, len=20), PIs1[2,], col = 2)

post1<-sim(m1, data = list(SL = seq(-3, 3, len=20), T = rep(22, 20))) # T = 22
means1<-apply(post1, 2, mean)
PIs1<-apply(post1, 2, PI)
lines(seq(-3, 3, len=20), means1, col = 3)
lines(seq(-3, 3, len=20), PIs1[1,], col = 3)
lines(seq(-3, 3, len=20), PIs1[2,], col = 3)





m1.2<- ulam(
    alist (
        K~dnorm(p, sigma), 
        logit(p) <- a[T] + b_SL[T] * SL + b_S[T] * S, 
        a[T]~dnorm(0,1), 
        b_SL[T]~dexp(1), 
        b_S[T]~dexp(1), 
        sigma~dexp(1)

    ), dat = list(K=d$K, SL=d$SL, T=d$T, S=d$S), chains=4, log_lik=TRUE
)
precis(m1.2, depth=2)

dashboard(m1.2)

# continuous number then sim??? no Logistic regression force you to use link

plot(SL, K)
# T = 2, S = 0.5
post1.2<-link(m1.2, data = list(SL = seq(-3, 3, len=20), S = rep(0.5, 20), T = rep(2, 20)))
means1.2<-apply(post1.2, 2, mean)
PI1.2<-apply(post1.2, 2, PI)
lines(seq(-3, 3, len=20), means1.2, col = 2)
lines(seq(-3, 3, len=20), PI1.2[1,], col = 2)
lines(seq(-3, 3, len=20), PI1.2[2,], col = 2)

# T = 22, S = 0.5
post1.2<-link(m1.2, data = list(SL = seq(-3, 3, len=20), S = rep(0.5, 20), T = rep(22, 20)))
means1.2<-apply(post1.2, 2, mean)
PI1.2<-apply(post1.2, 2, PI)
lines(seq(-3, 3, len=20), means1.2, col = 3)
lines(seq(-3, 3, len=20), PI1.2[1,], col = 3)
lines(seq(-3, 3, len=20), PI1.2[2,], col = 3)

# Others as you want
post1.2<-link(m1.2, data = list(SL = seq(-3, 3, len=20), S = rep(0.1, 20), T = rep(14, 20)))
means1.2<-apply(post1.2, 2, mean)
PI1.2<-apply(post1.2, 2, PI)
lines(seq(-3, 3, len=20), means1.2, col = 3)
lines(seq(-3, 3, len=20), PI1.2[1,], col = 3)
lines(seq(-3, 3, len=20), PI1.2[2,], col = 3)



m1.3<- ulam(
    alist (
        K~dnorm(p, sigma), 
        logit(p) <- a[T] + b_SL[T] * SL + b_S[T] * S + a_dK * dK, 
        a[T]~dnorm(0,1), 
        b_SL[T]~dexp(1), 
        b_S[T]~dexp(1), 
        a_dK~ dnorm(-1, 0.5), # this is not pooled since it should be same over time
        sigma~dexp(1)

    ), dat = list(K=d$K, SL=d$SL, T=d$T, S=d$S, dK=d$dK), chains=4, log_lik=TRUE
)
dashboard(m1.3)
precis(m1.3)


plot(S, K)
post1.3<-link(m1.3, data = list(S = seq(0, 1, len=20), SL = rep(0, 20), T = rep(4, 20), dK=rep(0.5, 20)))
means1.3<-apply(post1.3, 2, mean)
PI1.3<-apply(post1.3, 2, PI)
lines(seq(0, 1, len=20), means1.3, col = 3)
lines(seq(0, 1, len=20), PI1.3[1,], col = 3)
lines(seq(0, 1, len=20), PI1.3[2,], col = 3)


m1.4<- ulam(
    alist (
        K~dnorm(p, sigma), 
        logit(p) <- a[T] + b_SL[T] * SL + b_S[T] * S + a_dK * dK + U[T], 
        a[T]~dnorm(0,1), 
        b_SL[T]~dexp(1), 
        b_S[T]~dexp(1), 
        a_dK~ dnorm(-1, 0.5), # this is not pooled since it should be same over time
        U[T]~ dnorm(0, 1),
        sigma~dexp(1)

    ), dat = list(K=d$K, SL=d$SL, T=d$T, S=d$S, dK=d$dK), chains=4, log_lik=TRUE
)
dashboard(m1.4)
precis(m1.4)

plot(SL, K)
post1.4<-link(m1.4, data = list(SL = seq(-3, 3, len=20), S = rep(0.5, 20), T = rep(4, 20), dK=rep(0.5, 20)))
means1.4<-apply(post1.4, 2, mean)
PI1.4<-apply(post1.4, 2, PI)
lines(seq(-3, 3, len=20), means1.4, col = 3)
lines(seq(-3, 3, len=20), PI1.4[1,], col = 3)
lines(seq(-3, 3, len=20), PI1.4[2,], col = 3)


# We have not yet modeled this -> we have to do this !!!
plot(T, K)
post1.4<-link(m1.4, data = list(SL = rep(0, 25), S = rep(0.5, 25), T = 1:25, dK=rep(0.5, 25)))
means1.4<-apply(post1.4, 2, mean)
PI1.4<-apply(post1.4, 2, PI)

for (i in 1:25){
    points(i, means1.4[i], col = 2, lwd = 3)
    points(i, PI1.4[1,i], col = 2, lwd = 2)
    points(i, PI1.4[2,i], col = 2, lwd = 2)

}

# Not stratified model 
m1.5<- ulam(
    alist (
        K~dnorm(p, sigma), 
        logit(p) <- a + b_SL * SL + b_S * S + a_dK * dK + U, 
        a~dnorm(0,1), 
        b_SL~dexp(1), 
        b_S~dexp(1), 
        a_dK~ dnorm(-1, 0.5), # this is not pooled since it should be same over time
        U~ dnorm(0, 1),
        sigma~dexp(1)

    ), dat = list(K=d$K, SL=d$SL, T=d$T, S=d$S, dK=d$dK), chains=4, log_lik=TRUE
)
dashboard(m1.5)
precis(m1.5)

plot(SL, K)
post1.5<-link(m1.5, data = list(SL = seq(-3, 3, len=20), S = rep(0.5, 20), dK=rep(0.5, 20)))
means1.5<-apply(post1.5, 2, mean)
PI1.5<-apply(post1.5, 2, PI)
lines(seq(-3, 3, len=20), means1.5, col = 3)
lines(seq(-3, 3, len=20), PI1.5[1,], col = 3)
lines(seq(-3, 3, len=20), PI1.5[2,], col = 3)

plot(S, K)
post1.5<-link(m1.5, data = list(S = seq(0, 1, len=20), SL = rep(0, 20), dK=rep(0.5, 20)))
means1.5<-apply(post1.5, 2, mean)
PI1.5<-apply(post1.5, 2, PI)
lines(seq(-3, 3, len=20), means1.5, col = 3)
lines(seq(-3, 3, len=20), PI1.5[1,], col = 3)
lines(seq(-3, 3, len=20), PI1.5[2,], col = 3)

# Two models 
m2<- ulam(
    alist (
        K~dnorm(p, sigma), 
        logit(p) <- a + b_SL * SL + b_S * S + b_dK * dK + U, 
        a~dnorm(0,1), 
        b_SL~dexp(1), 
        b_S~dexp(1), 
        b_dK~ dnorm(-1, 0.5), # this is not pooled since it should be same over time
        sigma~dexp(1),

        S~dnorm(p3, sigma3),
        logit(p3)<- a_S[T] + b_dS * dS,
        b_dS~ dnorm(-1, 0.5), 
        a_S[T]~dnorm(0, 1),
        sigma3~dexp(1),

        U~ dnorm(a_T[T], sigma2),
        a_T[T]~dnorm(0, 1),
        sigma2~dexp(1)

    ), dat = list(K=d$K, SL=d$SL, T=d$T, S=d$S, dK=d$dK, dS=d$dS), chains=4, log_lik=TRUE
)

dashboard(m2)
precis(m2)

curve(dpois(5, x), xlim= c(0, 10))

# model the T by unknown Y
m2<- ulam(
    alist (
        K~dnorm(p, sigma), 
        logit(p) <- a + b_SL * SL + b_S * S + b_dK * dK + U, 
        a~dnorm(0,1), 
        b_SL~dexp(1), 
        b_S~dexp(1), 
        b_dK~ dnorm(-1, 0.5), # this is not pooled since it should be same over time
        sigma~dexp(1),

        S~dnorm(p3, sigma3),
        logit(p3)<- a_S[T] + b_dS * dS,
        b_dS~ dnorm(-1, 0.5), 
        a_S[T]~dnorm(0, 1),
        sigma3~dexp(1),

        U~ dnorm(mu2, sigma2),
        mu2<- a_T 
        a_T[T]~dnorm(0, 1),
        sigma2~dexp(1), 

        T ~ poisson(l),
        l<-a_lT + b_Y * Y, 
        a_lT~dnorm(0,1), 
        b_Y~ dnorm(0,1),
        Y~dunif(0, 100)


    ), dat = list(K=d$K, SL=d$SL, T=d$T, S=d$S, dK=d$dK, dS=d$dS), chains=4, log_lik=TRUE
)

dashboard(m2)
precis(m2)


rm(list=ls())

```


```{r tdm2}

T<-1:25
SL<-NULL
for (i in T)SL[i]<- rnorm(1, - (T[i] - 50)^2 * 2/ 800+ 3, 0.6)

T<- round(runif(200, 1,25))
SL<-NULL
for (i in 1:length(T)) SL[i]<- rnorm(1, - (1/ 40 )*(T[i] - 12.5)^2 + 2, 0.5)
plot(T, SL)

dS<-rbeta(length(T), 1.2,1)
S<-NULL
for(i in 1:length(T))S[i] <- rnorm(1, - 0.01*(T[i]-13) ^2 -2*dS[i] + 15, 1)/17
plot(density(S))
plot(dS, S)
curve((-2.2 * x + 0.01*mean(T)+ 5)/7.5, add=TRUE)
plot(T, S)
plot(T, dS)


dK<-rbeta(length(T), 1.4,1)
K<-NULL
for(i in 1:length(T))K[i] <- rnorm(1, 4*SL[i] + 0.01* (T[i]-13)^2 + 75* S[i]- 25*dK[i]+15, 1)/90
plot(density(K))
plot(dK, K)
plot(SL, K)
plot(T, K)
plot(S, K)
plot(dS, K)

d<-list(T=T, SL=SL, dS=dS, S=S, K=K, dK=dK)
d


```


Sea Level: https://www.sci.kagoshima-u.ac.jp/oyo/advanced/image/endo2015-3.jpg

```{r m1}
library("rethinking")



plot(S, SL)

m1<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, 
        a[T]~dnorm(0,1), 

        b_S[T]~dexp(1), 
        b_SL[T]~dexp(1), 
        b_dK~dexp(1), 
        U~dexp(1),

         S~dnorm(mu2, sigma),
         mu2 <- a_mu2[T] - b_mu2 * dS,
         a_mu2[T]~dnorm(0, 1),
         b_mu2~dexp(1),
         sigma~dexp(1)

    ), dat = d, chains=4, log_lik=TRUE, 
    #custom_block = my_stan_function,
)


plot(SL, S)
precis(lm(SL~S))

m2<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, # regularization
        a[T]~dnorm(0,1), 

        transpars>vector[25]:b_S<<-v[,1],
        transpars>vector[25]:b_SL<<-v[,2],

        matrix[25, 2]:v ~multi_normal_cholesky(c(a_bar1, a_bar2), Rho, sigma),
        a_bar1~dexp(1),
        a_bar2~dexp(1),# Sl should be -4 
        corr_matrix[2]:Rho ~lkj_corr(2),# this is not strong correlation, huh? 
        vector[2]: sigma ~dexp(1),

        b_dK~dexp(1), 
        U~dexp(1),

         S~dnorm(mu2, sigma2),
         mu2 <- a_mu2[T] - b_mu2 * dS,
         a_mu2[T]~dnorm(0, 1),
         b_mu2~dexp(1),
         sigma2~dexp(1)

    ), dat = d, chains=4, log_lik=TRUE, 
    #custom_block = my_stan_function,
)


precis(m2, depth=2)
dashboard(m2)
# Plots

plot(SL, K)
points(SL[d$T==2], K[d$T==2], col = 2, lwd= 3)
curve(0.16 + 0.05 * x + 0.87 * 0.2 - 0.29 * 0.5, add=TRUE)

# Posterior Prediction
post1<-sim(m2, data = list(SL = seq(-3, 3, len=20), T = rep(2, 20), S = rep(0.1, 20), dK = rep(0.5, 20), dS = rep(0.5, 20))) # T = 2 
means1<-apply(post1, 2, mean)
PIs1<-apply(post1, 2, PI)
lines(seq(-3, 3, len=20), means1, col = 2)
lines(seq(-3, 3, len=20), PIs1[1,], col = 2)
lines(seq(-3, 3, len=20), PIs1[2,], col = 2)

post1<-sim(m1, data = list(SL = seq(-3, 3, len=20), T = rep(22, 20))) # T = 22
means1<-apply(post1, 2, mean)
PIs1<-apply(post1, 2, PI)
lines(seq(-3, 3, len=20), means1, col = 3)
lines(seq(-3, 3, len=20), PIs1[1,], col = 3)
lines(seq(-3, 3, len=20), PIs1[2,], col = 3)


m3<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, # regularization
        a[T]~dnorm(0,1), 

        save>vector[25]:b_S<<- b_S_bar + b_S_sig * v[,1],
        save>vector[25]:b_SL<<- b_SL_bar + b_SL_sig * v[,2],

        matrix[25, 2]:v ~multi_normal_cholesky(a_bar, Rho, sigma),
        vector[2]: a_bar~dnorm(0, 1),
        corr_matrix[2]:Rho ~lkj_corr(2),# this is not strong correlation, huh? 
        vector[2]: sigma ~dexp(1),
        c(b_S_bar, b_SL_bar)~dexp(0,1),
        c(b_S_sig, b_SL_sig)~dexp(1),

        b_dK~dexp(1), 
        U~dexp(1),

         S~dnorm(mu2, sigma2),
         mu2 <- a_mu2[T] - b_mu2 * dS,
         a_mu2[T]~dnorm(0, 1),
         b_mu2~dexp(1),
         sigma2~dexp(1)

    ), dat = d, chains=4, cores=4, log_lik=TRUE, 
    #custom_block = my_stan_function,
)
precis(m3, depth=2)
dashboard(m3)

plot(SL, K)
points(SL[d$T==2], K[d$T==2], col = 2, lwd= 3)
curve(0.19 + 0.05 * x + 0.83 * 1 - 0.29 * 1, add=TRUE)

# Posterior Prediction
post1<-sim(m3, data = 
PIs1<-apply(post1, 2, PI)
lines(seq(-3, 3, len=20), means1, col = 2)
lines(seq(-3, 3, len=20), PIs1[1,], col = 2)
lines(seq(-3, 3, len=20), PIs1[2,], col = 2)


rm(list = ls())
```


Same things with actual number before normalize. 
and normalize within the model 

```{r d2}

T<-1:25
SL<-NULL
for (i in T)SL[i]<- rnorm(1, - (T[i] - 50)^2 * 2/ 800+ 3, 0.6)

T<- round(runif(200, 1,25)) 
SL<-NULL # m
for (i in 1:length(T)) SL[i]<- rnorm(1, - (1/ 40 )*(T[i] - 12.5)^2 + 2, 0.5)
plot(T, SL)

dS<-rbeta(length(T), 1.2,1) * 200 # km
S<-NULL
for(i in 1:length(T))S[i] <- rnorm(1, - 0.05*(T[i]-13) ^2 -dS[i]/100 + 15, 1)/17
plot(density(S))
plot(dS, S)
points(dS[T==2], S[T==2], col = 2, lwd =3)
points(dS[T==13], S[T==13], col = 3, lwd =3)
points(dS[T==7], S[T==7], col = 4, lwd =3)
plot(T, S)
plot(T, dS)


dK<-rbeta(length(T), 1.4,1)*400
K<-NULL
for(i in 1:length(T))K[i] <- rnorm(1, 4*SL[i] + 0.01* (T[i]-13)^2 + 75* S[i]- dK[i]/16, 1)/90
plot(density(K))
plot(dK, K)
plot(SL, K)
plot(T, K)
plot(S, K)
plot(dS, K)

d<-list(T=T, SL=SL, dS=dS, S=S, K=K, dK=dK)
d

```
# Model 3

- Model K with SL, S, dK by stratifying by T
- Correlation between SL and S 
- Model S with dS 

# Posterior distribution 


```{r m3}

# Is there correlation??/ Yes it is!!!!
plot(S, SL)
points(S[T==2], SL[T==2], col = 2, lwd=3)
points(S[T==3], SL[T==3], col = 3, lwd=3)

m3<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, # regularization

        vector[25]: a<-a_bar + siga * za[T],
        za[T]~dnorm(0,1),
        a_bar~dnorm(0,1),
        siga ~ dexp(1),

        transpars>vector[25]:b_S<<-v[,1],
        transpars>vector[25]:b_SL<<-v[,2],


        matrix[25, 2]:v ~multi_normal(c(SL_bar, S_bar), Rho, sigma),
        SL_bar~dnorm(0.12, 0.1),
        S_bar~dnorm(1.4, 0.5),
        corr_matrix[2]:Rho ~lkj_corr(2),# this is not strong correlation, huh? 
        vector[2]: sigma ~dexp(1),

        b_dK~dexp(1), # don't need regularization
        U~dexp(1),

        S~dnorm(mu2, sigma2),
        mu2 <- a2[T] - b_dS * dS,

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        za2[T]~dnorm(0,1),
        a2_bar~dnorm(0,1),
        siga2 ~ dexp(1),

        b_dS~dexp(1),

        sigma2~dexp(1)

    ), dat = d, chains=4, cores=4, log_lik=TRUE, 
    #custom_block = my_stan_function,
)

precis(m3, depth=2)
dashboard(m3)

plot(SL, K)
points(SL[d$T==2], K[d$T==2], col = 2, lwd= 3)
curve(-0.07 + 0.57* (-0.10)+ 3.03 * x + 0.07 * 0.2 - 0.03 * 100, add=TRUE)

# Posterior Check

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=2 )
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m3)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation


# Prediction model
means<-NULL
PIs<-NULL
means$a <- apply( post$a , 2 , mean )
means$b_SL <- apply( post$b_SL , 2 , mean )
means$b_S <- apply( post$b_S , 2 , mean )
means$b_dK <- apply( post$b_dK , 2 , mean )

plot(SL, K)
curve(means$a + means$b_SL * x + means$b_S * 0.3 - means$b_dK * 100, add = TRUE, col = 2)
PIs$a <- apply( post$a , 2 , PI )
PIs$b_SL <- apply( post$b_SL , 2 , PI )
PIs$b_S <- apply( post$b_S , 2 , PI )
PIs$b_dK <- apply( post$b_dK , 2 , PI )
curve(PIs$a[1] + PIs$b_SL[1] * x + PIs$b_S[1] * 0.3 - PIs$b_dK[1] * 100, add = TRUE, lty=2, col = 2)
curve(PIs$a[2] + PIs$b_SL[2] * x + PIs$b_S[2] * 0.3 - PIs$b_dK[2] * 100, add = TRUE, lty=2, col = 2)

curve(means$a + means$b_SL * x + means$b_S * 0.3 - means$b_dK * 200, add = TRUE, col = 3)
curve(PIs$a[1] + PIs$b_SL[1] * x + PIs$b_S[1] * 0.3 - PIs$b_dK[1] * 200, add = TRUE, lty=2, col = 3)
curve(PIs$a[2] + PIs$b_SL[2] * x + PIs$b_S[2] * 0.3 - PIs$b_dK[2] * 200, add = TRUE, lty=2, col = 3)
curve(means$a + means$b_SL * x + means$b_S * 0.3 - means$b_dK * 300, add = TRUE, col = 4)
curve(PIs$a[1] + PIs$b_SL[1] * x + PIs$b_S[1] * 0.3 - PIs$b_dK[1] * 300, add = TRUE, lty=2, col = 4)
curve(PIs$a[2] + PIs$b_SL[2] * x + PIs$b_S[2] * 0.3 - PIs$b_dK[2] * 300, add = TRUE, lty=2, col = 4)


plot(S, K)
curve(means$a + means$b_SL * 0 + means$b_S * x - means$b_dK * 100, add = TRUE, col = 2)
curve(PIs$a[1] + PIs$b_SL[1] * 0 + PIs$b_S[1] * x - PIs$b_dK[1] * 100, add = TRUE, lty=2, col = 2)
curve(PIs$a[2] + PIs$b_SL[2] * 0 + PIs$b_S[2] * x - PIs$b_dK[2] * 100, add = TRUE, lty=2, col = 2)

curve(means$a + means$b_SL * 0 + means$b_S * x - means$b_dK * 200, add = TRUE, col = 3)
curve(PIs$a[1] + PIs$b_SL[1] * 0 + PIs$b_S[1] * x - PIs$b_dK[1] * 200, add = TRUE, lty=2, col = 3)
curve(PIs$a[2] + PIs$b_SL[2] * 0 + PIs$b_S[2] * x - PIs$b_dK[2] * 200, add = TRUE, lty=2, col = 3)
curve(means$a + means$b_SL * 0 + means$b_S * x - means$b_dK * 300, add = TRUE, col = 4)
curve(PIs$a[1] + PIs$b_SL[1] * 0 + PIs$b_S[1] * x - PIs$b_dK[1] * 300, add = TRUE, lty=2, col = 4)
curve(PIs$a[2] + PIs$b_SL[2] * 0 + PIs$b_S[2] * x - PIs$b_dK[2] * 300, add = TRUE, lty=2, col = 4)


```

# Model4 
- Model K with SL, S, dK wby stratifying by T
- Correlation between SL and S 
- Model S with dS and y
- Model SL with Y 
- Y is normal? uniform? 



```{r m4}

16000 - 2400
# starting from -80
# 16000 to 10000 -> +30/5000 = 3/500
# 10000 to 7000 -> + 60 / 3000 = 10/500
# 7000 to 6000 -> + 4
# 6000 to 3000 -> - 5 / 3000 = -1 / 500
# 3000 to 2400 -> -5 / 500


curve((x +16)* 30/ 6 - 80, from = -16, to = -10, col = 2, xlim=c(-16.000, -2.400), ylim=c(-85, 10))
curve((x + 10) * 60/ 3 - 50, from = -10, to = -7.3, col = 2, add=TRUE)
lines(x=c(-7.3, -6), y = c(4, 4), col = 2,  add=TRUE) # -7.3 to 6
curve((x + 6) * (-5)/ 3 + 4, from = -6, to = -3, col = 2, add=TRUE)
curve((x + 3) * (-5)/ 0.5 - 1, from = -3, to = -2.4, col = 2, add=TRUE)

m4<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, # regularization

        vector[25]: a<-a_bar + siga * za[T],
        za[T]~dnorm(0,1),
        a_bar~dnorm(0,1),
        siga ~ dexp(1),

        transpars>vector[25]:b_S<<-v[,1],
        transpars>vector[25]:b_SL<<-v[,2],


        matrix[25, 2]:v ~multi_normal(c(SL_bar, S_bar), Rho, sigma),
        SL_bar~dnorm(0.12, 0.1),
        S_bar~dnorm(1.4, 0.5),
        corr_matrix[2]:Rho ~lkj_corr(2),# this is not strong correlation, huh? 
        vector[2]: sigma ~dexp(1),

        b_dK~dexp(1), # don't need regularization
        U~dexp(1),

        # Model S 
        S~dnorm(muS, sigmaS),
        muS <- a2[T] - b_dS * dS + b_Y[T] * Y[T],

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        za2[T]~dnorm(0,1),
        a2_bar~dnorm(0,1),
        siga2 ~ dexp(1),

        b_dS~dexp(1),
        vector[25]: b_Y<-b_Y_bar + sigy * zy[T],
        b_Y_bar~dnorm(0, 1),
        sigy~dexp(1),
        zy[T]~dnorm(0,1),

        sigmaS~dexp(1), 

        # Model SL 
        SL | Y[T]< -10.000 ~dnorm((Y[T] +16)* 30/ 6 - 80, sigmaSL),
        SL | Y[T]< -7.300 ~dnorm((Y[T] + 10) * 60/ 3 - 50, sigmaSL),
        SL | Y[T]< -6.000~dnorm(4, sigmaSL),
        SL | Y[T]< -3.000 ~dnorm((Y[T] + 6) * (-5)/ 3 + 4, sigmaSL),
        SL | Y[T]< -2.400 ~dnorm((Y[T] + 3) * (-5)/ 0.5 - 1, sigmaSL),
        sigmaSL ~ dexp(1),

        # Model Y
        vector[25]: Y~dunif(2.400, 16.000)


    ), dat = d, chains=4, cores=4, log_lik=TRUE, 
#    custom_block = my_stan_function
)
precis(m4, depth=2)
dashboard(m4)

# Posterior Check

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=2 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m4)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation


# Prediction model
means<-NULL
PIs<-NULL
means$a <- apply( post$a , 2 , mean )
means$b_SL <- apply( post$b_SL , 2 , mean )

means$b_S <- apply( post$b_S , 2 , mean )
means$b_dK <- apply( post$b_dK , 2 , mean )

plot(SL, K)
curve(means$a + means$b_SL * x + means$b_S * 0.3 - means$b_dK * 100, add = TRUE, col = 2)
PIs$a <- apply( post$a , 2 , PI )
PIs$b_SL <- apply( post$b_SL , 2 , PI )
PIs$b_S <- apply( post$b_S , 2 , PI )
PIs$b_dK <- apply( post$b_dK , 2 , PI )
curve(PIs$a[1] + PIs$b_SL[1] * x + PIs$b_S[1] * 0.3 - PIs$b_dK[1] * 100, add = TRUE, lty=2, col = 2)
curve(PIs$a[2] + PIs$b_SL[2] * x + PIs$b_S[2] * 0.3 - PIs$b_dK[2] * 100, add = TRUE, lty=2, col = 2)

curve(means$a + means$b_SL * x + means$b_S * 0.3 - means$b_dK * 200, add = TRUE, col = 3)
curve(PIs$a[1] + PIs$b_SL[1] * x + PIs$b_S[1] * 0.3 - PIs$b_dK[1] * 200, add = TRUE, lty=2, col = 3)
curve(PIs$a[2] + PIs$b_SL[2] * x + PIs$b_S[2] * 0.3 - PIs$b_dK[2] * 200, add = TRUE, lty=2, col = 3)
curve(means$a + means$b_SL * x + means$b_S * 0.3 - means$b_dK * 300, add = TRUE, col = 4)
curve(PIs$a[1] + PIs$b_SL[1] * x + PIs$b_S[1] * 0.3 - PIs$b_dK[1] * 300, add = TRUE, lty=2, col = 4)
curve(PIs$a[2] + PIs$b_SL[2] * x + PIs$b_S[2] * 0.3 - PIs$b_dK[2] * 300, add = TRUE, lty=2, col = 4)


plot(S, K)
curve(means$a + means$b_SL * 0 + means$b_S * x - means$b_dK * 100, add = TRUE, col = 2)
curve(PIs$a[1] + PIs$b_SL[1] * 0 + PIs$b_S[1] * x - PIs$b_dK[1] * 100, add = TRUE, lty=2, col = 2)
curve(PIs$a[2] + PIs$b_SL[2] * 0 + PIs$b_S[2] * x - PIs$b_dK[2] * 100, add = TRUE, lty=2, col = 2)

curve(means$a + means$b_SL * 0 + means$b_S * x - means$b_dK * 200, add = TRUE, col = 3)
curve(PIs$a[1] + PIs$b_SL[1] * 0 + PIs$b_S[1] * x - PIs$b_dK[1] * 200, add = TRUE, lty=2, col = 3)
curve(PIs$a[2] + PIs$b_SL[2] * 0 + PIs$b_S[2] * x - PIs$b_dK[2] * 200, add = TRUE, lty=2, col = 3)
curve(means$a + means$b_SL * 0 + means$b_S * x - means$b_dK * 300, add = TRUE, col = 4)
curve(PIs$a[1] + PIs$b_SL[1] * 0 + PIs$b_S[1] * x - PIs$b_dK[1] * 300, add = TRUE, lty=2, col = 4)
curve(PIs$a[2] + PIs$b_SL[2] * 0 + PIs$b_S[2] * x - PIs$b_dK[2] * 300, add = TRUE, lty=2, col = 4)

rm(list=ls())

```


# Standardized data

```{r d2}

curve((x +16)* 30/ 6 - 80, from = -16, to = -10, col = 2, xlim=c(-16.000, -2.400), ylim=c(-85, 10))
curve((x + 10) * 60/ 3 - 50, from = -10, to = -7.3, col = 2, add=TRUE)
lines(x=c(-7.3, -6), y = c(4, 4), col = 2,  add=TRUE) # -7.3 to 6
curve((x + 6) * (-5)/ 3 + 4, from = -6, to = -3, col = 2, add=TRUE)
curve((x + 3) * (-5)/ 0.5 - 1, from = -3, to = -2.4, col = 2, add=TRUE)


T<- round(runif(200, 1,25)) 
SL<-NULL # m
for (i in 1:length(T)) SL[i]<- rnorm(1, - (1/ 40 )*(T[i] - 12.5)^2 + 2, 0.5)
plot(T, SL)

dS<-rbeta(length(T), 1.2,1) * 200 # km
S<-NULL
for(i in 1:length(T))S[i] <- rnorm(1, - 0.05*(T[i]-13) ^2 -dS[i]/100 + 15, 1)/17
plot(density(S))
plot(dS, S)
points(dS[T==2], S[T==2], col = 2, lwd =3)
points(dS[T==13], S[T==13], col = 3, lwd =3)
points(dS[T==7], S[T==7], col = 4, lwd =3)
plot(T, S)
plot(T, dS)


dK<-rbeta(length(T), 1.4,1)*400
K<-NULL
for(i in 1:length(T))K[i] <- rnorm(1, 4*SL[i] + 0.01* (T[i]-13)^2 + 75* S[i]- dK[i]/16, 1)/90


# Standardize


length(standardize(SL))
length((T - mean(T))/sd(T))


d2<-list(T=T, SL=standardize(SL), dS=standardize(dS), S=standardize(S), K=standardize(K), dK=standardize(dK))
d2

plot(density(d2$K))
plot(d2$dK, d2$K)
plot(d2$SL, d2$K)
plot(d2$T, d2$K)
plot(d2$S, d2$K)
plot(d2$dS, d2$K)
plot(d2$dS, d2$S)


```

```{r m5}
m5<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, # regularization

        vector[25]: a<-a_bar + siga * za[T],
        za[T]~dnorm(0,1),
        a_bar~dnorm(-0.5,0.5),
        siga ~ dexp(1),

        transpars>vector[25]:b_S<<-v[,1],
        transpars>vector[25]:b_SL<<-v[,2],


        matrix[25, 2]:v ~multi_normal(c(SL_bar, S_bar), Rho, sigma),
        SL_bar~dnorm(0.12, 0.1),
        S_bar~dnorm(1.4, 0.5),
        corr_matrix[2]:Rho ~lkj_corr(2),# this is not strong correlation, huh? 
        vector[2]: sigma ~dexp(1),

        b_dK~dexp(1), # don't need regularization
        U~dexp(1),

        # Model S 
        S~dnorm(muS, sigmaS),
        muS <- a2[T] - b_dS * dS + b_Y[T] * Y[T],

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        za2[T]~dnorm(0,1),
        a2_bar~dnorm(0,0.1),
        siga2 ~ dexp(1),

        b_dS~dexp(1),
        vector[25]: b_Y<-b_Y_bar + sigy * zy[T],
        b_Y_bar~dnorm(0, 1),
        sigy~dexp(1),
        zy[T]~dnorm(0,1),

        sigmaS~dexp(1), 

        # Model SL 
        SL | Y[T]< -10.000 ~dnorm((Y[T] +16)* 30/ 6 - 80, sigmaSL),
        SL | Y[T]< -7.300 ~dnorm((Y[T] + 10) * 60/ 3 - 50, sigmaSL),
        SL | Y[T]< -6.000~dnorm(4, sigmaSL),
        SL | Y[T]< -3.000 ~dnorm((Y[T] + 6) * (-5)/ 3 + 4, sigmaSL),
        SL | Y[T]< -2.400 ~dnorm((Y[T] + 3) * (-5)/ 0.5 - 1, sigmaSL),
        sigmaSL ~ dexp(1),

        # Model Y
        vector[25]: Y~dunif(2.400, 16.000)


    ), dat = d2, chains=4, cores=4, log_lik=TRUE, 
#    custom_block = my_stan_function
)
dashboard(m5)
precis(m5, depth=2)

# Posterior Check

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=2 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m5)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation

# Prediction model
means<-NULL
PIs<-NULL
means$a <- apply( post$a , 2 , mean )
means$b_SL <- apply( post$b_SL , 2 , mean )

means$b_S <- apply( post$b_S , 2 , mean )
means$b_dK <- apply( post$b_dK , 2 , mean )

plot(d2$SL, d2$K)
points(d2$SL[T==2], d2$K[T==2], col = 2, lwd=3)
PIs$a <- apply( post$a , 2 , PI )
PIs$b_SL <- apply( post$b_SL , 2 , PI )
PIs$b_S <- apply( post$b_S , 2 , PI )
PIs$b_dK <- apply( post$b_dK , 2 , PI )
curve(means$a[2] + means$b_SL[2] * x + means$b_S[2] * mean(d2$S[T==2]) - means$b_dK * (mean(d2$dK[T==2])), add = TRUE, col = 2)
curve(PIs$a[1,2] + PIs$b_SL[1,2] * x + PIs$b_S[1,2] * (mean(d2$S[T==2])) - PIs$b_dK[1] * (mean(d2$dK[T==2])), add = TRUE, lty=2, col = 2)
curve(PIs$a[2,2] + PIs$b_SL[2,2] * x + PIs$b_S[2,2] * (mean(d2$S[T==2])) - PIs$b_dK[2] * (mean(d2$dK[T==2])), add = TRUE, lty=2, col = 2)


# Messy Plot
for(i in 1:25){
points(d2$SL[T==i], d2$K[T==i], col = i, lwd=3)
curve(means$a[i] + means$b_SL[i] * x + means$b_S[i] * mean(d2$S[T==i]) - means$b_dK * mean(d2$dK[T==i]), add = TRUE, col = i)
curve(PIs$a[1,i] + PIs$b_SL[1,i] * x + PIs$b_S[1,i] * mean(d2$S[T==i]) - PIs$b_dK[1,] * mean(d2$dK[T==i]), add = TRUE, lty=2, col = i)
curve(PIs$a[2,i] + PIs$b_SL[2,i] * x + PIs$b_S[2,i] * mean(d2$S[T==i]) - PIs$b_dK[2,] * mean(d2$dK[T==i]), add = TRUE, lty=2, col = i)

}


plot(d2$S, d2$K)
points(d2$S[T==2], d2$K[T==2], col =2, lwd = 2)
curve(means$a[2] + means$b_SL[2] * mean(d2$SL[T==2]) + means$b_S[2] * x - means$b_dK * mean(d2$dK), add = TRUE, col = 2)
curve(PIs$a[1,2] + PIs$b_SL[1,2] * mean(d2$SL[T==2]) + PIs$b_S[1,2] * x - PIs$b_dK[1] *  mean(d2$dK), add = TRUE, lty=2, col = 2)
curve(PIs$a[2,2] + PIs$b_SL[2,2] * mean(d2$SL[T==2]) + PIs$b_S[2,2] * x - PIs$b_dK[2] *  mean(d2$dK), add = TRUE, lty=2, col = 2)



# do not feed in the array in the curve function 
a<-0:100
curve(a*x)
curve(a*x, xlim=c(0, 10))
curve(a*x, xlim=c(0, 20))
a<-1:10
curve(a*x)



rm(list=ls())
```


```{r m6}
library("rethinking")
m6<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S + b_Y[T] * Y[T] - b_dK * dK, # regularization

        vector[25]: a<-a_bar + siga * za[T],
        za[T]~dnorm(0,1),
        a_bar~dnorm(-0.5,0.5),
        siga ~ dexp(1),

        vector[25]: b_Y<-b_Y_bar + sigb_Y * zb_Y[T],
        zb_Y[T]~dnorm(0,1),
        b_Y_bar~dnorm(0.5,0.5),
        sigb_Y ~ dexp(1),

        transpars>vector[25]:b_S<<-v[,1],
        transpars>vector[25]:b_SL<<-v[,2],


        matrix[25, 2]:v ~multi_normal(c(SL_bar, S_bar), Rho, sigma),
        SL_bar~dnorm(0.12, 0.1),
        S_bar~dnorm(1.4, 0.5),
        corr_matrix[2]:Rho ~lkj_corr(2),# this is not strong correlation, huh? 
        vector[2]: sigma ~dexp(1),


        b_dK~dexp(1), # don't need regularization
        U~dexp(1),

        # Model S 
        S~dnorm(muS, sigmaS),
        muS <- a2[T] - b_dS * dS + b_Y_S[T] * Y[T],

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        za2[T]~dnorm(0,1),
        a2_bar~dnorm(0,0.1),
        siga2 ~ dexp(1),

        b_dS~dexp(1),
        vector[25]: b_Y_S<-b_Y_S_bar + sigys * zys[T],
        b_Y_S_bar~dnorm(0, 1),
        sigys~dexp(1),
        zys[T]~dnorm(0,1),

        sigmaS~dexp(1), 

        # Model SL 
        SL | Y[T] < -16.000 ~dnorm(- 80, sigmaSL),
        SL | Y[T] > -16.000 && Y[T]< -10.000 ~dnorm((Y[T] +16)* 30/ 6 - 80, sigmaSL),
        SL | Y[T] > -10.000 && Y[T]< -7.300 ~dnorm((Y[T] + 10) * 60/ 3 - 50, sigmaSL),
        SL | Y[T] > -7.300 && Y[T]< -6.000~dnorm(4, sigmaSL),
        SL | Y[T] > -6.000 && Y[T]< -3.000 ~dnorm((Y[T] + 6) * (-5)/ 3 + 4, sigmaSL),
        SL | Y[T] > -3.000 && Y[T]< -2.400 ~dnorm((Y[T] + 3) * (-5)/ 0.5 - 1, sigmaSL),
        # SL | Y[T] > -16.000 && Y[T]< -10.000 ~dnorm((-13 +16)* 30/ 6 - 80, sigmaSL),
        # SL | Y[T] > -10.000 && Y[T]< -7.300 ~dnorm((-8.6 + 10) * 60/ 3 - 50, sigmaSL),
        # SL | Y[T] > -7.300 && Y[T]< -6.000~dnorm(4, sigmaSL),
        # SL | Y[T] > -6.000 && Y[T]< -3.000 ~dnorm((-4.5 + 6) * (-5)/ 3 + 4, sigmaSL),
        # SL | Y[T] > -3.000 && Y[T]< -2.400 ~dnorm((-2.7 + 3) * (-5)/ 0.5 - 1, sigmaSL),
        SL | Y[T] > -2.400 ~dnorm(0, sigmaSL),
        sigmaSL ~ dexp(10),

       # Model Y scale it to 0 to 1 
        vector[25]: Y~dunif(-16.000, -2.400)


    ), dat = d2, chains=4, cores=4, log_lik=TRUE, 
#    custom_block = my_stan_function
)
dashboard(m6)
precis(m6, depth=2)

curve(dexp(x,8))
# Posterior Check

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=3 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m6)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation

# Prediction model
means<-NULL
PIs<-NULL
means$a <- apply( post$a , 2 , mean )
means$b_SL <- apply( post$b_SL , 2 , mean )

means$b_S <- apply( post$b_S , 2 , mean )
means$b_dK <- apply( post$b_dK , 2 , mean )

means$b_Y <- apply( post$b_Y , 2 , mean )
means$Y <- apply( post$Y , 2 , mean )

PIs$a <- apply( post$a , 2 , PI )
PIs$b_SL <- apply( post$b_SL , 2 , PI )
PIs$b_S <- apply( post$b_S , 2 , PI )
PIs$b_dK <- apply( post$b_dK , 2 , PI )
PIs$b_Y <- apply( post$b_Y , 2 , PI )
PIs$Y <- apply( post$Y , 2 , PI )

plot(d2$SL, d2$K)
points(d2$SL[T==2], d2$K[T==2], col = 2, lwd=3)
curve(means$a[2] + means$b_SL[2] * x + means$b_S[2] * mean(d2$S[T==2]) + means$b_Y[2] * means$Y[2] - means$b_dK * (mean(d2$dK[T==2])), add = TRUE, col = 2)
curve(PIs$a[1,2] + PIs$b_SL[1,2] * x + PIs$b_S[1,2] * (mean(d2$S[T==2])) + PIs$b_Y[1,2] * PIs$Y[1,2] - PIs$b_dK[1] * (mean(d2$dK[T==2])), add = TRUE, lty=2, col = 2)
curve(PIs$a[2,2] + PIs$b_SL[2,2] * x + PIs$b_S[2,2] * (mean(d2$S[T==2])) + PIs$b_Y[2,2] * PIs$Y[2,2] - PIs$b_dK[2] * (mean(d2$dK[T==2])), add = TRUE, lty=2, col = 2)


# Messy Plot
for(i in 1:25){
points(d2$SL[T==i], d2$K[T==i], col = i, lwd=3)
curve(means$a[i] + means$b_SL[i] * x + means$b_S[i] * mean(d2$S[T==i]) + means$b_Y[2] * means$Y[2] - means$b_dK * mean(d2$dK[T==i]), add = TRUE, col = i)
curve(PIs$a[1,i] + PIs$b_SL[1,i] * x + PIs$b_S[1,i] * mean(d2$S[T==i]) + PIs$b_Y[1,2] * PIs$Y[1,2] - PIs$b_dK[1,] * mean(d2$dK[T==i]), add = TRUE, lty=2, col = i)
curve(PIs$a[2,i] + PIs$b_SL[2,i] * x + PIs$b_S[2,i] * mean(d2$S[T==i]) + PIs$b_Y[2,2] * PIs$Y[2,2] - PIs$b_dK[2,] * mean(d2$dK[T==i]), add = TRUE, lty=2, col = i)

}


plot(d2$S, d2$K)
points(d2$S[T==2], d2$K[T==2], col =2, lwd = 2)
curve(means$a[2] + means$b_SL[2] * mean(d2$SL[T==2]) + means$b_S[2] * x + means$b_Y[2] * means$Y[2] - means$b_dK * mean(d2$dK), add = TRUE, col = 2)
curve(PIs$a[1,2] + PIs$b_SL[1,2] * mean(d2$SL[T==2]) + PIs$b_S[1,2] * x + PIs$b_Y[1,2] * PIs$Y[1,2] - PIs$b_dK[1] *  mean(d2$dK), add = TRUE, lty=2, col = 2)
curve(PIs$a[2,2] + PIs$b_SL[2,2] * mean(d2$SL[T==2]) + PIs$b_S[2,2] * x + PIs$b_Y[2,2] * PIs$Y[1,2] - PIs$b_dK[2] *  mean(d2$dK), add = TRUE, lty=2, col = 2)




# using sim?
n<-2000

length(post$a[,1])
dam<-rnorm(n, 
        mean = post$a[,1] + post$b_SL[,1] * d2$SL[T==1] + post$b_S[,1] * mean(d2$S[T==1]) + post$b_Y[,1] * post$Y[,1] - post$b_dK[1] * mean(d2$dK[T==1]), 
        sd = post$U)
mean <- post$a[,1] + post$b_SL[,1] * rep(mean(d2$SL[T==1]), 2000) + post$b_S[,1] * ran + post$b_Y[,1] * post$Y[,1] - post$b_dK[1] * rep(mean(d2$dK), 2000)
length(dam)
dens(dam)
plot(ran, mean)
plot(ran, dam)
simK<-matrix(NA, nrow = 25, ncol = n) 
for(i in 1:25){
    # I just put d2$variable[T==i] but might want to be mean()
    simK[i,]<-rnorm(n, # n == length(post)
        mean = post$a[,i] + post$b_SL[,i] * mean(d2$SL[T==i]) + post$b_S[,i] * mean(d2$S[T==i]) + post$b_Y[,i] * post$Y[,i] - post$b_dK[i] * mean(d2$dK[T==i]), 
        sd = post$U)
}

length(simK[2])

apply(matrix(simK[1,]), 2, PI )
apply(matrix(simK[2,]), 2, PI )
apply(matrix(simK[3,]), 2, PI )
apply(matrix(simK[4,]), 2, PI )
sim_PI <- apply(simK, 1, PI )
sim_PI <- t(sim_PI)

length(sim_PI)
sim_PI[1,1]- sim_PI[1,2]
sim_PI[2,1]- sim_PI[2,2]
sim_PI[3,1]- sim_PI[3,2]

plot(d2$T, d2$K)
points(1:25, sim_PI[,1],col=2, lwd = 3)
points(1:25, sim_PI[,2], col=2, lwd =3)

for(i in 1:25)lines(c(i,i), sim_PI[i,1:2], col =2, lwd = 2)



# This does not work
dens(post$b_S)
dens(post$b_SL)
dens(d2$SL)
Sseq<-seq(-3, 3, len = 100)
mu<-sim(m6,data=list(SL=Sseq, T = rep(0, 100), S = rep(1, 100), dK = rep(1, 100), dS = rep(1, 100)))



# for SL when T ==2
simK<-matrix(NA, nrow = 20, ncol = n) 
range(SL)
sSL<-seq(-2.5, 3, len=20)
sSL[1]
for(j in 1:20){
    # I just put d2$variable[T==i] but might want to be mean()
    i <-2
    simK[j,]<-rnorm(n, # n == length(post)
        mean = post$a[,i] + post$b_SL[,i] * sSL[j] + post$b_S[,i] * mean(d2$S[T==i]) + post$b_Y[,i] * post$Y[,i] - post$b_dK[i] * mean(d2$dK[T==i]), 
        sd = post$U)
}

length(simK[,1])

sim_PI <- apply(simK, 1, PI )
sim_PI <- t(sim_PI)

plot(d2$SL, d2$K)
lines(sSL, sim_PI[,1],col=2, lwd = 3, lty=2)
lines(sSL, sim_PI[,2],col=2, lwd = 3, lty=2)


# everything10
simK<-matrix(NA, nrow = 100, ncol = n*25) 
sSL<-seq(-2.5, 3, len=100)
for(j in 1:100){
    # I just put d2$variable[T==i] but might want to be mean()    
    simK[j,]<-rnorm(n*25, # n == length(post)
        mean = array(post$a) + array(post$b_SL) * sSL[j] + array(post$b_S) * mean(d2$S) + array(post$b_Y) * array(post$Y) - array(rep(post$b_dK,25)) * mean(d2$dK), 
        sd = post$U)
}
sim_PI <- apply(simK, 1, PI )
sim_PI <- t(sim_PI)


plot(d2$SL, d2$K)
lines(sSL, sim_PI[,1],col=2, lwd = 3, lty=2)
lines(sSL, sim_PI[,2],col=2, lwd = 3, lty=2)


rm(list=ls())
```


# This does not goes well, adjust the initial data by adding Y (hidden)

```{r d3}
library("rethinking")
library(dplyr)
myFunc <- function(x) {
case_when(
    -16.000 <= x && x <= -10   ~ (x + 16) * 30 / 6 - 80,
    -10.000 < x  && x <= -7.300 ~ (x + 10) * 60 / 3 - 50,
    -7.300 < x   && x <= -6    ~ 4,
    -6.000 < x   && x <= -3.000 ~ (x + 6) * (-5) / 3 + 4,
    -3.000 < x   && x <= -2.400 ~ (x + 3) * (-5) / 0.5 - 1,
    TRUE                      ~ 1 # This acts as your final 'else'
  )

}
curve(myFunc(x), col = 2, xlim=c(-16.000, -2.400), ylim=c(-85, 10))

Y<- runif(200, -16, -2.4)
T<- NULL
SL<-NULL # m
for (i in 1:length(Y)) {
    T[i]<-rpois(1,  (Y[i]+16)/13.6 * 25)
    if(T[i] < 1) {T[i]<- 1}
    else if(T[i] > 25) {T[i]<- 25}
    SL[i]<- rnorm(1, myFunc(Y[i]), 5)
}
plot(T, SL)
plot(Y, T)
plot(Y, SL)

dS<-rbeta(length(T), 1.2,1) * 200 # km
S<-NULL
for(i in 1:length(T))S[i] <- rnorm(1, 0.3*Y[i] -dS[i]/50, 0.1)/10 + 1
range(S)
plot(density(S))
plot(dS, S)
points(dS[T==2], S[T==2], col = 2, lwd =4)
points(dS[T==13], S[T==13], col = 3, lwd =4)
points(dS[T==7], S[T==7], col = 4, lwd =4)
plot(T, S)
plot(Y, S)
plot(T, dS)


dK<-rbeta(length(T), 1.4,1)*400
K<-NULL
for(i in 1:length(T)){
    K[i] <- rnorm(1, 0.2*SL[i] + 0.02* Y[i] + 0.2* S[i]- dK[i]/50, 0.1)/45 + 0.8
    K[i]<- K[i] / (K[i] + S[i])
    S[i]<- S[i] / (K[i] + S[i])
    }
range(K)
plot(density(K))
plot(dK, K)
plot(SL, K)
plot(T, K)
plot(Y, K)
plot(S, K)
points(S[T==2], K[T==2], col = 2, lwd =4)
points(S[T==13], K[T==13], col = 3, lwd =4)
points(S[T==7], K[T==7], col = 4, lwd =4)
plot(dS, K)

# Standardize
length(standardize(SL))
length((T - mean(T))/sd(T))


d3<-list(T=T, SL=standardize(SL), dS=standardize(dS), S=S, K=K, dK=standardize(dK))
#d2

plot(density(d3$K))
plot(d3$dK, d3$K)
plot(d3$SL, d3$K)
plot(d3$T, d3$K)
plot(d3$S, d3$K)
plot(d3$dS, d3$K)
plot(d3$dS, d3$S)
plot(Y, d3$S)


```

```{r m7}
library("rethinking")

plot(d3$S, d3$SL)
points(d3$S[T==2], d3$SL[T==2], col = 2, lwd = 4)
points(d3$S[T==3], d3$SL[T==3], col = 3, lwd = 4)
points(d3$S[T==4], d3$SL[T==4], col = 5, lwd = 4)
m7<- ulam(
    alist (
        K~normal(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S + b_Y[T] * Y[T] - b_dK * dK, # regularization

        # We do not know anything about this. 
        # mean should start at 0 and has small variation 
        vector[25]: a<-a_bar + siga * za[T],
        a_bar~normal(0,0.5),
        siga ~ exponential(1),
        za[T]~normal(0,0.3),

        # In the same T, we do not know how Y changes, although it would be close to each other 
        vector[25]: b_Y<-b_Y_bar + sigb_Y * zb_Y[T],
        b_Y_bar~normal(0,0.5),
        sigb_Y ~ exponential(1),
        zb_Y[T]~normal(0,0.3),


        # We want to get SL and S positively correlated to K, 
        # but we do not know for sure without data. 
        # this is not too strong correlation, huh? 
        transpars>vector[25]:b_S<<-v[,1],
        transpars>vector[25]:b_SL<<-v[,2],

        matrix[25, 2]:v ~multi_normal(c(SL_bar, S_bar), Rho, sigma),
        SL_bar~normal(0, 0.5),
        S_bar~normal(0, 0.5),
        corr_matrix[2]:Rho ~lkj_corr(5),
        vector[2]: sigma ~exponential(1),

        # don't need regularization
        # We are hoping/believing b_dK is positive 
        b_dK~normal(0, 0.5), 
        U~exponential(1),

        # Model S 
        S~normal(muS, sigmaS),
        muS <- a2[T] - b_dS * dS + b_Y_S[T] * Y[T],

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        a2_bar~normal(0,0.5), 
        siga2 ~ exponential(1),
        za2[T]~normal(0,0.3),

        # We are hoping/believing b_dK is positive 
        b_dS~normal(0,0.5),

        vector[25]: b_Y_S<-b_Y_S_bar + sigys * zys[T],
        b_Y_S_bar~normal(0, 0.5),
        sigys~exponential(1),
        zys[T]~normal(0,0.3),

        sigmaS~exponential(1), 

        # Model SL 
        SL | Y[T] < -16.000 ~normal(- 80, sigmaSL),
        SL | Y[T] >= -16.000 && Y[T]<= -10.000 ~normal((Y[T] +16)* 30/ 6 - 80, sigmaSL),
        SL | Y[T] > -10.000 && Y[T]<= -7.300 ~normal((Y[T] + 10) * 60/ 3 - 50, sigmaSL),
        SL | Y[T] > -7.300 && Y[T]<= -6.000~normal(4, sigmaSL),
        SL | Y[T] > -6.000 && Y[T]<= -3.000 ~normal((Y[T] + 6) * (-5)/ 3 + 4, sigmaSL),
        SL | Y[T] > -3.000 && Y[T]<= -2.400 ~normal((Y[T] + 3) * (-5)/ 0.5 - 1, sigmaSL),
        # SL | Y[T] > -16.000 && Y[T]< -10.000 ~dnorm((-13 +16)* 30/ 6 - 80, sigmaSL),
        # SL | Y[T] > -10.000 && Y[T]< -7.300 ~dnorm((-8.6 + 10) * 60/ 3 - 50, sigmaSL),
        # SL | Y[T] > -7.300 && Y[T]< -6.000~dnorm(4, sigmaSL),
        # SL | Y[T] > -6.000 && Y[T]< -3.000 ~dnorm((-4.5 + 6) * (-5)/ 3 + 4, sigmaSL),
        # SL | Y[T] > -3.000 && Y[T]< -2.400 ~dnorm((-2.7 + 3) * (-5)/ 0.5 - 1, sigmaSL),
        SL | Y[T] > -2.400 ~normal(0, sigmaSL),
        sigmaSL ~ exponential(10),

# CHECKKKK
       # Model Y scale it to 0 to 1 
        vector[25]: Y~uniform(-16.000, -2.400)


    ), dat = d3, chains=4, cores=4, log_lik=TRUE, 
#    custom_block = my_stan_function
)
dashboard(m7)
precis(m7, depth=2)

# Posterior Check

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=3 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m7)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation

# Prediction model
means<-NULL
PIs<-NULL
means$a <- apply( post$a , 2 , mean )
means$b_SL <- apply( post$b_SL , 2 , mean )

means$b_S <- apply( post$b_S , 2 , mean )
means$b_dK <- apply( post$b_dK , 2 , mean )

means$b_Y <- apply( post$b_Y , 2 , mean )
means$Y <- apply( post$Y , 2 , mean )

PIs$a <- apply( post$a , 2 , PI )
PIs$b_SL <- apply( post$b_SL , 2 , PI )
PIs$b_S <- apply( post$b_S , 2 , PI )
PIs$b_dK <- apply( post$b_dK , 2 , PI )
PIs$b_Y <- apply( post$b_Y , 2 , PI )
PIs$Y <- apply( post$Y , 2 , PI )

plot(d3$SL, d3$K)
points(d3$SL[T==2], d3$K[T==2], col = 2, lwd=3)
curve(means$a[2] + means$b_SL[2] * x + means$b_S[2] * mean(d3$S[T==2]) + means$b_Y[2] * means$Y[2] - means$b_dK * (mean(d3$dK[T==2])), add = TRUE, col = 2)
curve(PIs$a[1,2] + PIs$b_SL[1,2] * x + PIs$b_S[1,2] * (mean(d3$S[T==2])) + PIs$b_Y[1,2] * PIs$Y[1,2] - PIs$b_dK[1] * (mean(d3$dK[T==2])), add = TRUE, lty=2, col = 2)
curve(PIs$a[2,2] + PIs$b_SL[2,2] * x + PIs$b_S[2,2] * (mean(d3$S[T==2])) + PIs$b_Y[2,2] * PIs$Y[2,2] - PIs$b_dK[2] * (mean(d3$dK[T==2])), add = TRUE, lty=2, col = 2)


# Messy Plot
for(i in 1:25){
points(d3$SL[T==i], d3$K[T==i], col = i, lwd=3)
curve(means$a[i] + means$b_SL[i] * x + means$b_S[i] * mean(d3$S[T==i]) + means$b_Y[2] * means$Y[2] - means$b_dK * mean(d3$dK[T==i]), add = TRUE, col = i)
curve(PIs$a[1,i] + PIs$b_SL[1,i] * x + PIs$b_S[1,i] * mean(d3$S[T==i]) + PIs$b_Y[1,2] * PIs$Y[1,2] - PIs$b_dK[1,] * mean(d3$dK[T==i]), add = TRUE, lty=2, col = i)
curve(PIs$a[2,i] + PIs$b_SL[2,i] * x + PIs$b_S[2,i] * mean(d3$S[T==i]) + PIs$b_Y[2,2] * PIs$Y[2,2] - PIs$b_dK[2,] * mean(d3$dK[T==i]), add = TRUE, lty=2, col = i)

}


plot(d3$S, d3$K)
points(d3$S[T==2], d3$K[T==2], col =2, lwd = 2)
curve(means$a[2] + means$b_SL[2] * mean(d3$SL[T==2]) + means$b_S[2] * x + means$b_Y[2] * means$Y[2] - means$b_dK * mean(d3$dK), add = TRUE, col = 2)
curve(PIs$a[1,2] + PIs$b_SL[1,2] * mean(d3$SL[T==2]) + PIs$b_S[1,2] * x + PIs$b_Y[1,2] * PIs$Y[1,2] - PIs$b_dK[1] *  mean(d3$dK), add = TRUE, lty=2, col = 2)
curve(PIs$a[2,2] + PIs$b_SL[2,2] * mean(d3$SL[T==2]) + PIs$b_S[2,2] * x + PIs$b_Y[2,2] * PIs$Y[1,2] - PIs$b_dK[2] *  mean(d3$dK), add = TRUE, lty=2, col = 2)




# using sim?
n<-2000

length(post$a[,1])
dam<-rnorm(n, 
        mean = post$a[,1] + post$b_SL[,1] * d2$SL[T==1] + post$b_S[,1] * mean(d2$S[T==1]) + post$b_Y[,1] * post$Y[,1] - post$b_dK[1] * mean(d2$dK[T==1]), 
        sd = post$U)
mean <- post$a[,1] + post$b_SL[,1] * rep(mean(d2$SL[T==1]), 2000) + post$b_S[,1] * ran + post$b_Y[,1] * post$Y[,1] - post$b_dK[1] * rep(mean(d2$dK), 2000)
length(dam)
dens(dam)
plot(ran, mean)
plot(ran, dam)
simK<-matrix(NA, nrow = 25, ncol = n) 
for(i in 1:25){
    # I just put d2$variable[T==i] but might want to be mean()
    simK[i,]<-rnorm(n, # n == length(post)
        mean = post$a[,i] + post$b_SL[,i] * mean(d2$SL[T==i]) + post$b_S[,i] * mean(d2$S[T==i]) + post$b_Y[,i] * post$Y[,i] - post$b_dK[i] * mean(d2$dK[T==i]), 
        sd = post$U)
}

length(simK[2])

apply(matrix(simK[1,]), 2, PI )
apply(matrix(simK[2,]), 2, PI )
apply(matrix(simK[3,]), 2, PI )
apply(matrix(simK[4,]), 2, PI )
sim_PI <- apply(simK, 1, PI )
sim_PI <- t(sim_PI)

length(sim_PI)
sim_PI[1,1]- sim_PI[1,2]
sim_PI[2,1]- sim_PI[2,2]
sim_PI[3,1]- sim_PI[3,2]

plot(d2$T, d2$K)
points(1:25, sim_PI[,1],col=2, lwd = 3)
points(1:25, sim_PI[,2], col=2, lwd =3)

for(i in 1:25)lines(c(i,i), sim_PI[i,1:2], col =2, lwd = 2)



# This does not work
dens(post$b_S)
dens(post$b_SL)
dens(d2$SL)
Sseq<-seq(-3, 3, len = 100)
mu<-sim(m6,data=list(SL=Sseq, T = rep(0, 100), S = rep(1, 100), dK = rep(1, 100), dS = rep(1, 100)))



# for SL when T ==2
simK<-matrix(NA, nrow = 20, ncol = n) 
range(SL)
sSL<-seq(-2.5, 3, len=20)
sSL[1]
for(j in 1:20){
    # I just put d2$variable[T==i] but might want to be mean()
    i <-2
    simK[j,]<-rnorm(n, # n == length(post)
        mean = post$a[,i] + post$b_SL[,i] * sSL[j] + post$b_S[,i] * mean(d2$S[T==i]) + post$b_Y[,i] * post$Y[,i] - post$b_dK[i] * mean(d2$dK[T==i]), 
        sd = post$U)
}

length(simK[,1])

sim_PI <- apply(simK, 1, PI )
sim_PI <- t(sim_PI)

plot(d2$SL, d2$K)
lines(sSL, sim_PI[,1],col=2, lwd = 3, lty=2)
lines(sSL, sim_PI[,2],col=2, lwd = 3, lty=2)


# everything10
simK<-matrix(NA, nrow = 100, ncol = n*25) 
sSL<-seq(-2.5, 3, len=100)
for(j in 1:100){
    # I just put d2$variable[T==i] but might want to be mean()    
    simK[j,]<-rnorm(n*25, # n == length(post)
        mean = array(post$a) + array(post$b_SL) * sSL[j] + array(post$b_S) * mean(d2$S) + array(post$b_Y) * array(post$Y) - array(rep(post$b_dK,25)) * mean(d2$dK), 
        sd = post$U)
}
sim_PI <- apply(simK, 1, PI )
sim_PI <- t(sim_PI)


plot(d2$SL, d2$K)
lines(sSL, sim_PI[,1],col=2, lwd = 3, lty=2)
lines(sSL, sim_PI[,2],col=2, lwd = 3, lty=2)


rm(list=ls())
```
# 7 does not work as it tries to model Y 

without Y and no correlation, but does not hide SL
```{r m8}
library("rethinking")

plot(d3$S, d3$SL)
points(d3$S[T==2], d3$SL[T==2], col = 2, lwd = 4)
points(d3$S[T==3], d3$SL[T==3], col = 3, lwd = 4)
points(d3$S[T==4], d3$SL[T==4], col = 5, lwd = 4)
m8<- ulam(
    alist (
        K~normal(mu, U), 
        logit(mu) <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, 

        # We do not know anything about this. 
        # mean should start at 0 and has small variation 
        vector[25]: a<-a_bar + siga * za[T],
        a_bar~normal(0,0.3),
        siga ~ exponential(1),
        za[T]~normal(0,0.3),

        # We want to get SL and S positively correlated to K due to Y 
        # but we do not know for sure without data. 
        # this is not too strong correlation, huh? 
        # transpars>vector[25]:b_S<<-v[,1],
        # transpars>vector[25]:b_SL<<-v[,2],

        # matrix[25, 2]:v ~multi_normal(c(SL_bar, S_bar), Rho, sigma),
        # SL_bar~normal(0, 0.2),
        # S_bar~normal(0, 0.2),
        # corr_matrix[2]:Rho ~lkj_corr(1),
        # vector[2]: sigma ~exponential(1),

        vector [25]: b_S~normal(0, 0.2),
        vector [25]: b_SL~normal(0, 0.2),

        # don't need regularization
        # We are hoping/believing b_dK is positive 
        b_dK~normal(0, 0.5),

        U~exponential(1),

        # Model S 
        S~normal(muS, sigmaS),
        logit(muS) <- a2[T] - b_dS * dS,

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        a2_bar~normal(0,0.3), 
        siga2 ~ exponential(1),
        za2[T]~normal(0,0.3),

        # We are hoping/believing b_dS is positive 
        b_dS~normal(0,0.5),

        sigmaS~exponential(1)


    ), dat = d3, chains=4, cores=4, log_lik=TRUE, 
#    custom_block = my_stan_function
)
dashboard(m8)
precis(m8, depth=2)

# Posterior Check
post <- extract.samples(m8)

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=2 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m8)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation


# using sim? or Other stuffs   


plot(d3$SL, d3$K)
SLseq<-seq(-2, 2, len = 20)
KbySL<- sapply(SLseq, 
    function (i) inv_logit(post$a + post$b_SL * i + post$b_S * mean(d3$S)))
mean(KbySL[,1])
PIs<- apply( KbySL , 2 , PI )
means<- apply( KbySL , 2 , mean )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
curve(inv_logit(mean(post$a_bar) + mean(post$b_SL) * x + mean(post$b_S) * mean(d3$S)), add=TRUE)

b_SL[T] * SL + b_S[T] * S - b_dK * dK

plot(d3$S, d3$K)
Sseq<-seq(0, 0.75, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$a + post$b_S * i+ ))
PIs<- apply( KbyS , 2 , PI )
means<- apply( KbyS , 2 , mean )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
curve(inv_logit(mean(post$a_bar) + mean(post$b_S) * x), add=TRUE)

plot(d3$dK, d3$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$a - rep(post$b_dK, 25) * i + post$b_S * mean(d3$S)))
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d3$dS, d3$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$a2 - rep(post$b_dS, 25) * i ))
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


rm(list=ls())
```


But we do not know SL from the data 

one way: add SL for each T 


another way: add Y 

# One 

```{r m9}
library("rethinking")

d4<- list(K=d3$K, T=d3$T, S=d3$S, dK=d3$dK, dS=d3$dS)

plot(d4$T, d3$SL)
curve(inv_logit(x - 10) + 10,xlim = c(-1, 10) )

m9<- ulam(
    alist (
        K~normal(mu, U), 
        logit(mu) <- a[T] + b_SL[T] * SL[T] + b_S[T] * S - b_dK * dK,

        # We do not know anything about this. 
        # mean should start at 0 and has small variation 
        vector[25]: a<-a_bar + siga * za[T],
        a_bar~normal(0,0.3),
        siga ~ exponential(1),
        za[T]~normal(0,0.3),


        # We want to get SL and S positively correlated to K due to Y 
        # but we do not know for sure without data. 
        # this is not too strong correlation, huh? 
        # transpars>vector[25]:b_S<<-v[,1],
        # transpars>vector[25]:b_SL<<-v[,2],

        # matrix[25, 2]:v ~multi_normal(c(SL_bar, S_bar), Rho, sigma),
        # SL_bar~normal(0, 0.2),
        # S_bar~normal(0, 0.2),
        # corr_matrix[2]:Rho ~lkj_corr(1),
        # vector[2]: sigma ~exponential(1),

        vector [25]: b_S~normal(0, 0.2),
        vector [25]: b_SL~normal(0, 0.2),


        # don't need regularization
        # We are hoping/believing b_dK is positive 
        b_dK~normal(0, 0.5),
        U~exponential(1),

        # Model S 
        S~normal(muS, sigmaS),
        logit(muS) <- a2[T] - b_dS * dS,

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        a2_bar~normal(0,0.3), 
        siga2 ~ exponential(1),
        za2[T]~normal(0,0.3),

        # We are hoping/believing b_dS is positive 
        b_dS~normal(0,0.5),

        sigmaS~exponential(1), 

        # Model SL 
        vector[25]:SL~normal( muSL , sigSL),
        vector[25]:logit(muSL)<- aSL + T * b2_SL,
        aSL~normal(-4, 0.1),
        b2_SL~normal(0.3, 0.1),


        sigSL~exponential(6)


    ), dat = d4, chains=4, cores=4, log_lik=TRUE, 
#    custom_block = my_stan_function
)
dashboard(m9)
precis(m9, depth=2)



# Posterior Check
post <- extract.samples(m9)

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=2 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m9)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation


# using sim? or Other stuffs   

# Scale SL between 0 and 1
SL<-(d3$SL - min(d3$SL))/(max(d3$SL) - min(d3$SL))

plot(SL, d3$K)
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, 
    function (i) inv_logit(mean(post$a) + post$b_SL * i + post$b_S * mean(d3$S)))
mean(KbySL[,1])
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
curve(mean(post$a_bar) + mean(post$b_SL) * x + mean(post$b_S) * mean(d3$S), add=TRUE)


plot(d3$S, d3$K)
Sseq<-seq(0, 1, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$a + post$b_S * i))
PIs<- apply( KbyS , 2 , PI )
means<- apply( KbyS , 2 , mean )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(d3$dK, d3$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$a - rep(post$b_dK, 25) * i + post$b_S * mean(d3$S)))
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d3$dS, d3$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$a2 - rep(post$b_dS, 25) * i ))
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

# How about T and SL
plot(d3$T, SL)
dSseq<-1:25
SLbyT<- sapply(dSseq, 
    function (i) inv_logit(post$aSL + post$b2_SL * i ))
head(post$aSL)
PIs<- apply( SLbyT , 2 , PI )
means<- apply( SLbyT , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

curve(inv_logit(0.3* x - 4), add = TRUE)

# Scale the SL from 0 to 1 


rm(list=ls())
```



# put the correlation information through y to hide SL 

Then there are some correlation between S and SL, not knowing if it's Y or something else
... this is ... Model 3 

(still have to hide SL tho) 

```{r m9}
library("rethinking")

plot(Y, d3$S)

m9<- ulam(
    alist (
        K~normal(mu, U), 
        logit(mu) <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK,

        # We do not know anything about this. 
        # mean should start at 0 and has small variation 
        vector[25]: a<-a_bar + siga * za[T],
        a_bar~normal(0,0.3),
        siga ~ exponential(1),
        za[T]~normal(0,0.3),


        # We want to get SL and S positively correlated to K due to Y 
        # but we do not know for sure without data. 
        # this is not too strong correlation, huh? 
        transpars>vector[25]:b_S<<-v[,1],
        transpars>vector[25]:b_SL<<-v[,2],

        matrix[25, 2]:v ~multi_normal(c(S_bar, SL_bar), Rho, sigma),
        S_bar~normal(0, 0.2),
        SL_bar~normal(0, 0.2),
        corr_matrix[2]:Rho ~lkj_corr(1),
        vector[2]: sigma ~exponential(1),


        # don't need regularization
        # We are hoping/believing b_dK is positive 
        b_dK~normal(0, 0.5),
        U~exponential(1),

        # Model S 
        S~normal(muS, sigmaS),
        logit(muS) <- a2[T] - b_dS * dS,

        vector[25]: a2<-a2_bar + siga2 * za2[T],
        a2_bar~normal(0,0.3), 
        siga2 ~ exponential(1),
        za2[T]~normal(0,0.3),

        # We are hoping/believing b_dS is positive 
        b_dS~normal(0,0.5),

        sigmaS~exponential(1)


    ), dat = d3, chains=4, cores=4, log_lik=TRUE, 
#    custom_block = my_stan_function
)
dashboard(m9)
precis(m9, depth=2)

# Posterior Check
post <- extract.samples(m9)

# Correlation?
R <- rlkjcorr( 1e4 , K=2 , eta=2 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))

post <- extract.samples(m9)
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation


# using sim? or Other stuffs   


plot(d3$SL, d3$K)
SLseq<-seq(-2, 2, len = 20)
KbySL<- sapply(SLseq, 
    function (i) inv_logit(mean(post$a) + post$b_SL * i + post$b_S * mean(d3$S)))
mean(KbySL[,1])
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
curve(mean(post$a_bar) + mean(post$b_SL) * x + mean(post$b_S) * mean(d3$S), add=TRUE)


plot(d3$S, d3$K)
Sseq<-seq(0, 1, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$a + post$b_S * i))
PIs<- apply( KbyS , 2 , PI )
means<- apply( KbyS , 2 , mean )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(d3$dK, d3$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$a - rep(post$b_dK, 25) * i + post$b_S * mean(d3$S)))
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d3$dS, d3$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$a2 - rep(post$b_dS, 25) * i ))
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


rm(list=ls())
```
⌈



```{r m32}
m32<- ulam(
    alist (
        K~dnorm(mu, U), 
        mu <- a[T] + b_SL[T] * SL + b_S[T] * S - b_dK * dK, # regularization
        a[T]~dnorm(0,1), 

        save>vector[25]:b_S<<- b_S_bar + b_S_sig * v[,1],
        save>vector[25]:b_SL<<- b_SL_bar + b_SL_sig * v[,2],

        matrix[25, 2]:v ~multi_normal_cholesky(a_bar, Rho, sigma),
        vector[2]: a_bar~dnorm(0, 1),
        corr_matrix[2]:Rho ~lkj_corr(2),# this is not strong correlation, huh? 
        vector[2]: sigma ~dexp(1),
        c(b_S_bar, b_SL_bar)~dexp(0,1),
        c(b_S_sig, b_SL_sig)~dexp(1),

        b_dK~dexp(1), 
        U~dexp(1),

         S~dnorm(mu2, sigma2),
         mu2 <- a_mu2[T] - b_mu2 * dS,
         a_mu2[T]~dnorm(0, 1),
         b_mu2~dexp(1),
         sigma2~dexp(1)

    ), dat = d, chains=4, cores=4, log_lik=TRUE, 
    #custom_block = my_stan_function,
)
precis(m32, depth=2)
dashboard(m32)

```



# No staratification
```{r m}
dum<-NULL
for(i in 1:100) dum[i]<-rpois(1, (runif(1, -16 + 2.4) + 16 - 2.4 / (-2.4 + 16)))
dum
dens(dum)

mT0<-ulam(
    alist(
         # Model T 
        T~poisson((Y + 16)/ ( -2.4 + 16) * 25),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )
    ), dat = d, chain = 4, core = 4, log_lik = TRUE
)

mT1<-ulam(
    alist(
         # Model T 
        T~normal((Y + 16)/ ( -2.4 + 16) * 25, 1),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )
    ), dat = d, chain = 4, core = 4, log_lik = TRUE
)


dashboard(mT0)
precis(mT0, depth = 2)
dashboard(mT1)
precis(mT1, depth = 2)

d$T
post <- extract.samples(mT0)
post1 <- extract.samples(mT1)
p<-NULL
p1<-NULL
for(i in 1:200) {
    p[i]<-mean(post$Y[,i])
    p1[i]<-mean(post1$Y[,i])
}

plot(T, Y)
points(T, p, col = 2)
points(T, p1, col = 3)

# The mT0 is the better one of two 
#

```

```{r reasoning}
curve(inv_logit(-0.78 + 0.39 * x + 0.72 * mean(d$S) + 0.06 * mean(post$Y) -0.23 * mean(d$dK)), add = TRUE, col = 3)

#curve(inv_logit(-0.62+0.5*x+0.35*mean(SL)+0.07*mean(Y)-0.23*mean(d$dK)), add = TRUE)


# Notice the curve is not as steep as overall data look like. 
# This should be due to the stratification. 
points(post$SL[1, d$T == 2], d$K[d$T == 2], col = 3, lwd = 4)
points(SL[d$T == 2], d$K[d$T == 2], col = 4, lwd = 4)

```

######################################################

This is Backup for Y between -16 to -2.4


# Model 1 

- Package Installation
```{r packages, error = FALSE, warning = FALSE, message = FALSE, results = "hide"}
library("rethinking")
library(dplyr)
library(dagitty)
```

## Causal effects of the DAG 
```{r DAG}
kdag<-dagitty("dag {
  S-> K <- T 
  T -> S
  T -> SL -> K
}")
adjustmentSets(kdag, exposure="SL", outcome = "K")# This shows the backdoor path and how to stratify

```


## Synthetic Data
- create a synthetic data with assumptions. See overview.pdf to check the data relationship. 
```{r data, fig.show="hold", out.width="50%"}

# Function of SL by Y 
myFunc <- function(x) {
case_when(
    -16.000 <= x & x <= -9   ~ (x + 16) * 60 / 7 - 80,
    -9.000 < x  & x <= -7.500 ~ (x + 9) * 46 / 3 - 20,
    -7.500 < x   & x <= -4.300    ~ 3,
    -4.300 < x   & x <= -3.000 ~ (x + 4.3) * (-60) / 13 + 3,
    -3.000 < x   & x <= -2.400 ~ (x + 3) * 5 / 3 ,
    TRUE                      ~ 1 
  )
}
curve(myFunc(x), col = 2, xlim=c(-16.000, -2.400), ylim=c(-85, 10))

# Initialize Y 
Y<- runif(200, -16, -2.4)

# T and SL caused by Y
T<- NULL
SL<-NULL
for (i in 1:length(Y)) {
    T[i]<-rpois(1,  (Y[i]+16)/13.6 * 25)
    if(T[i] < 1) {T[i]<- 1}
    else if(T[i] > 25) {T[i]<- 25} # Make sure T is between 1 and 25
    SL[i]<- rnorm(1, myFunc(Y[i]), 5) 
}

SL <- (SL - min(SL)) / (max(SL) - min(SL)) # to estimate easily, make it between 0 and 1 

# Check the data
plot(T, SL)
plot(Y, T)
plot(Y, SL)

# initialize dS with random numbers
dS<-rbeta(length(T), 1.2,1) * 200 
# S is affected by Y and dS (omit T as T and S are correlated by Y)
S<-NULL
for(i in 1:length(T))S[i] <- rnorm(1, 0.3*Y[i] -dS[i]/50, 0.1)/25 + 0.4

# Check the data
range(S) # try to make less than 0.5
plot(density(S))
plot(dS, S)
points(dS[T==2], S[T==2], col = 2, lwd =4)
points(dS[T==13], S[T==13], col = 3, lwd =4)
points(dS[T==7], S[T==7], col = 4, lwd =4)
plot(T, S)
plot(Y, S)
plot(T, dS) # no correlation 
plot(SL, S)

# Initialize dK with random numbers
dK<-rbeta(length(T), 1.4,1)*400
# K is affected by SL, Y, S, dK
K<-NULL
for(i in 1:length(T))K[i] <- rnorm(1, SL[i] + 0.1* Y[i] + S[i]- dK[i]/250, 0.2)/9 + 0.4

# Check the data
range(K)# try to make less than 0.5
plot(density(K))
plot(dK, K)
plot(SL, K)
plot(Y, K)
plot(S, K)
plot(dS, K)

# Standardize data without SL, S, K (those are ratio)
d<-list(T=T, dS=standardize(dS), S=S, K=K, dK=standardize(dK))

# check the data
plot(density(d$K))
plot(d$dK, d$K)
plot(SL, d$K)
plot(d$T, d$K)
plot(d$S, d$K)
plot(d$dS, d$S)
plot(Y, d$S)
plot(d$dS, d$K) # no correlation

```


## Prior Prediction 
- Also see overview.pdf to know statistical model and prior
- There are a few messages to warning my priors. 
- This would be due to the prior of Y, which is very weakly initialized. Since it only shows up during first iteration of warmup, we can safely ignore it. 
- If you really fix it, you can use normal distribution with large sigma and constraints between -16 and -2.4
```{r m1}
# Model 1 
m1<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_Y[T] * Y[i] + bK_dK * dK,

        # We do not know anything about this but think similar for all group. 
        # So use partial pooling to regularize the estimate. 
        # mean should start at 0 and has small variation.
        
        # In this case (m1), we do not include correlation. 
        # This is because multilevel models of S and SL includes the information of Y 
        # But we have to regularize. 
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_S <- bK_S_bar + sigma_bK_S * z_bK_S[T],
        vector [25]: bK_SL <- bK_SL_bar + sigma_bK_SL * z_bK_SL[T],
        vector [25]: bK_Y <- bK_Y_bar + sigma_bK_Y * z_bK_Y[T],
        c(aK_bar, bK_S_bar,  bK_SL_bar, bK_Y_bar) ~ normal(0,0.3),
        c(sigma_aK, sigma_bK_S, sigma_bK_SL, sigma_bK_Y) ~ exponential(6), # want low variation among groups
        z_aK[T] ~ normal(0,0.01),
        z_bK_S[T] ~ normal(0,0.01),
        z_bK_SL[T] ~ normal(0,0.01),
        z_bK_Y[T] ~ normal(0,0.01),

        # don't need regularization as dK is not affected by T
        # We are hoping/believing b_dK is negative, but this is not prior. 
        bK_dK ~ normal(0, 0.5),
        U ~ exponential(1),


    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_dS * dS + bS_Y[T] * Y[i],

        # I am hoping/believing b_dS is negative and b_Y_S is positive 
        # partial pooling for the same reason for a2 and b_Y_S
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],
        c(aS_bar, bS_Y_bar) ~ normal(0,0.1), 
        c(sigma_aS, sigmabS_Y) ~ exponential(6),
        z_aS[T] ~ normal(0,0.05),
        zbS_Y[T] ~ normal(0,0.05),

        # don't need regularization
        bS_dS ~ normal(0,0.1),
        sigma_S ~ exponential(1),


        # Model SL by Y 
        # The log function and prior was refered to the previous research of Sea Level Change over time. 
        # This function is abstruct so I will have to actually stan code to provide accurate estimates
        # ulam code in r does not allow us to hartd code the conditional function. 
        # vector[200]: SL ~ normal(muSL + aSL2, sigmaSL),
        vector[200]: SL ~ normal(mu_SL, sigma_SL),
        vector[200]: logit(mu_SL) <- aSL + Y[i] * bSL,
        aSL ~ normal(8, 0.01), # very low variation since this part is hard coded. 
        bSL ~ normal(0.8, 0.01), # very low variation since this part is hard coded. 

        sigma_SL ~ exponential(1), 

         # Model T, This should be hard coded based on the previous research
        T ~ poisson((Y + 16)/ ( -2.4 + 16) * 25),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1" # SL is between 0 and 1 
        #, T    = "lower=1,upper=25"  #this is unnecessary since T is provided in the dat 
    ),  chains=4, cores=4, log_lik=TRUE
)


dashboard(m1)
precis(m1, depth=1)
```

# Posterior Check
```{r post}
post <- extract.samples(m1)
```

## Hidden parameters 
- The hidden parameters are estimated by the model constraints and distributions. 
- It is expected that the difference is small and has similar trend as the original data.
```{r hidden, fig.show="hold", out.width="50%"}
# Posterior Y changes 
plot(1:200, Y)
points(1:200, post$Y[1,], col = 2)
for(i in 1 : 200) lines(c(i, i), c(post$Y[1,i], Y[i]), col = 2)

plot(Y, d$K)
points(post$Y[1,], d$K, col = 2)
for(i in 1 : 200) lines(c(post$Y[1,i], Y[i]), c(d$K[i], d$K[i]), col = 2)

plot(SL, d$K)
points(post$SL[1,], d$K, col = 2)
for(i in 1 : 200) lines(c(post$SL[1,i], SL[i]), c(d$K[i], d$K[i]), col = 2)
points(post$SL[2,], d$K, col = 3)
for(i in 1 : 200) lines(c(post$SL[2,i], SL[i]), c(d$K[i], d$K[i]), col = 3)

plot(Y, SL)
points(post$Y[1,], post$SL[1,], col=2)
for(i in 1 : 200) lines(c(Y[i], post$Y[1,i]), c(SL[i], post$SL[1,i]), col = 2)


plot(Y, d$T)
points(post$Y[1,], d$T ,col = 2)
for (i in 1:200) lines(c(Y[i], post$Y[1,i]), c(d$T[i],d$T[i]), col = 2)

```


=> this shows that the difference is not very large and hidden parameters have similar trend as the original (synthetized) data. 

## Posterior Simulation 

### Model K 
```{r postpredK, fig.show="hold", out.width="50%"}
plot(d$S, d$K)
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$aK + post$bK_S * i + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(SL, d$K)
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, # Tricky but t and t to calculate estimates by group, though dK does not have group so do not need 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + post$bK_SL * i + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d$dK, d$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * i )) 
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(Y, d$K)
Yseq<-seq(-16, -2.4, len = 20)
KbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + post$bK_Y * i + rep(post$bK_dK,25) * d$K )) 
PIs<- apply( KbyY , 2 , PI )
means<- apply( KbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
```

### Model S by dS and Y 
- By the way, there is no post-S exists as it seems absolutely same as post-SL
- would be because it is one of the predictors that i provided data 

```{r postpredS, fig.show="hold", out.width="50%"}
post$SL[1,] == post$S[1,] 

plot(d$dS, d$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * i + post$bS_Y * colMeans(post$Y))) # Used mean of each simulated data
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$S)
Yseq<-seq(-16, -2.4, len = 20)
SbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * d$dS + post$bS_Y * i))
means<- apply( SbyY , 2 , mean )
PIs<- apply( SbyY , 2 , PI )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
```

- Functions catch the trend very well!

### Model SL by Y 

```{r postpredSL, fig.show="hold", out.width="50%"}
plot(Y, SL)
points(post$Y[1,], post$SL[1,], col = 2)
Yseq<-seq(-16, -2.4, len = 20)
SLbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aSL + post$bSL * i))
PIs<- apply( SLbyY , 2 , PI )
means<- apply( SLbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

```


- The model 1 looks good. All posterior prediction plots shows inclusion of data points inside the model scope. 

- but also I am wondering if I use correlation between S and SL instead of modeling Y in K model. 
This is the m2


# Model 2 (m2)
- see priors in overview.pdf, but this include correlation between SL and S
- warnings are annoying but relatively few. 
```{r m2}
# Model 2
m2<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_Y[T] * Y[i] + bK_dK * dK,

        # We do not know anything about this but think similar for all group. 
        # So use partial pooling to regularize the estimate. 
        # mean should start at 0 and has small variation.
        
        # In this case (m2), we do include correlation between S and SL due to Y. 
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_Y <- bK_Y_bar + sigma_bK_Y * z_bK_Y[T],
        c(aK_bar, bK_Y_bar)~normal(0,0.3),
        c(sigma_aK, sigma_bK_Y) ~ exponential(6),
        z_aK[T]~normal(0,0.01),
        z_bK_Y[T] ~ normal(0,0.01),

        transpars>vector[25]:bK_S <<- v[,1],
        transpars>vector[25]:bK_SL <<- v[,2],

        matrix[25, 2]:v ~ multi_normal(c(bK_S_bar, bK_SL_bar), RhoSSL, sigmaSSL),
        c(bK_S_bar, bK_SL_bar) ~ normal(0, 0.3),
        corr_matrix[2]:RhoSSL ~ lkj_corr(1),# this is not too strong but good correlation
        vector[2]: sigmaSSL ~ exponential(2), 


        # don't need regularization
        # We are hoping/believing b_dK is negative, but this is not prior. 
        bK_dK ~ normal(0, 0.5),
        U ~ exponential(1),


    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_dS * dS,

        # I am hoping/believing b_dS is negative and b_Y_S is positive 
        # partial pooling for the same reason for a2 and b_Y_S
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],

        c(aS_bar, bS_Y_bar) ~ normal(0,0.3), 
        c(sigma_aS, sigmabS_Y) ~ exponential(6),
        z_aS[T] ~ normal(0,0.01),
        zbS_Y[T] ~ normal(0,0.01),

        # don't need regularization
        bS_dS ~ normal(0,0.1),
        sigma_S ~ exponential(1),


        # Model SL by Y 
        # The log function and prior was refered to the previous research of Sea Level Change over time. 
        # This function is abstruct so I will have to actually stan code to provide accurate estimates
        # ulam code in r does not allow us to hartd code the conditional function. 
        # vector[200]: SL ~ normal(muSL + aSL2, sigmaSL),
        vector[200]: SL ~ normal(mu_SL, sigma_SL),
        vector[200]: logit(mu_SL) <- aSL + Y[i] * bSL,
        aSL ~ normal(8, 0.01), # very low variation since this part is hard coded. 
        bSL ~ normal(0.8, 0.01), # very low variation since this part is hard coded. 

        sigma_SL ~ exponential(1), 

         # Model T 
        T ~ poisson((Y + 16)/ ( -2.4 + 16) * 25),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1" # SL is between 0 and 1 
    ),chains=4, cores=4, log_lik=TRUE 
)


dashboard(m2)
precis(m2, depth=1)
```

-> Looks worse sampling efficiency. 

# Posterior Check
```{r post2}
post <- extract.samples(m2)
```
- I do not check the Hidden parameters since I have never touched it since m1

## Correlation?
Rho value shows strong negative correlation between them. 
```{r post2cor, fig.show="hold", out.width="50%"}
R <- rlkjcorr( 1e4 , K=2 , eta=1 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation
```
## Posterior Predictions

```{r post2pred, fig.show="hold", out.width="50%"}
plot(d$S, d$K)
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$aK + post$bK_S * i + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(SL, d$K)
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, # Tricky but t and t to calculate estimates by group, though dK does not have group so do not need 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + post$bK_SL * i + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(d$dK, d$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * i )) 
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$K)
Yseq<-seq(-16, -2.4, len = 20)
KbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + post$bK_Y * i + rep(post$bK_dK,25) * d$K )) 
PIs<- apply( KbyY , 2 , PI )
means<- apply( KbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

# Model S
plot(d$dS, d$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * i )) # Used mean of each simulated data
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


```

K estimation by SL is a bit narrower than it supposed to be but other than that, everything looks great. 
Let me compare those two models 

```{r comp, fig.show="hold", out.width="50%"}
# Compare models
model_comparison <- compare(m1, m2)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison)
```

### Key Columns in the compare Output:
- WAIC / PSIS: Lower values indicate better out-of-sample predictive accuracy.
- dWAIC / dPSIS: The difference in information criteria between the current model and the top-ranked (best) model.
- dSE: If the standard error is large relative to the difference, the two models have statistically indistinguishable predictive capacities.
- weight: The Akaike weight, which represents the relative probability or evidence that a given model will make the best predictions among the set of compared models

Therefore, the m1 is better than the other one. 

# Model 3

```{r m3}
m3<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK + bK_S * S + bK_SL * SL[i] + bK_Y * Y[i] + bK_dK * dK + bK_T[T],

        # We do not know anything about this but think similar for all group. 
        # So use partial pooling to regularize the estimate. 
        # mean should start at 0 and has small variation.
        
        c(aK, bK_S,  bK_SL, bK_Y, bK_dK) ~ normal(0,0.3),
        vector[25]: bK_T <- bK_T_bar + sigma_bK_T * z_bK_T[T],
        bK_T_bar ~ normal(0, 0.3), # regularization
        sigma_bK_T ~ exponential(6), 
        z_bK_T[T] ~ normal(0, 1),

        U ~ exponential(1),


    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS + bS_dS * dS + bS_Y * Y[i] + bS_T[T],

        # I am hoping/believing b_dS is negative and b_Y_S is positive 
        # partial pooling for the same reason for a2 and b_Y_S
        c(aS, bS_Y, bS_dS) ~ normal(0,0.1), 
        vector[25]: bS_T <- bS_T_bar + sigma_bS_T * z_bS_T[T],
        bS_T_bar ~ normal(0, 0.001), # regularization
        sigma_bS_T ~ exponential(10), 
        z_bS_T[T] ~ normal(0, 0.1),

        sigma_S ~ exponential(1),

        # Model SL by Y 
        # The log function and prior was refered to the previous research of Sea Level Change over time. 
        # This function is abstruct so I will have to actually stan code to provide accurate estimates
        # ulam code in r does not allow us to hartd code the conditional function. 
        # vector[200]: SL ~ normal(muSL + aSL2, sigmaSL),
        vector[200]: SL ~ normal(mu_SL, sigma_SL),
        vector[200]: logit(mu_SL) <- aSL + Y[i] * bSL,
        aSL ~ normal(8, 0.01), # very low variation since this part is hard coded. 
        bSL ~ normal(0.8, 0.01), # very low variation since this part is hard coded. 

        sigma_SL ~ exponential(1), 

         # Model T 
        T ~ poisson((Y + 16)/ ( -2.4 + 16) * 25),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1" # SL is between 0 and 1 
    ),chains=4, cores=4, log_lik=TRUE
)

dashboard(m3)
precis(m3, depth=1)

```

-> Looks worse sampling efficiency. 

# Posterior Check
```{r post3}
post <- extract.samples(m3)
```
- Again, I do not check the Hidden parameters since I have never touched it since m1
```{r post3check, fig.show="hold", out.width="50%"}

# Model K

plot(d$S, d$K)
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$aK + post$bK_S * i + post$bK_SL * colMeans(post$SL) + post$bK_Y * colMeans(post$Y)  + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] ))
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(SL, d$K)
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq,  # This is tricky but use bK_T for T value and others are not stratified so randomly assigned
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * i + post$bK_Y * colMeans(post$Y)  + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] ))
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d$dK, d$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_Y * colMeans(post$Y)  + post$bK_dK * i + colMeans(post$bK_T)[d$T] ))
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$K)
Yseq<-seq(-16, -2.4, len = 20)
KbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_Y * i + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] ))
PIs<- apply( KbyY , 2 , PI )
means<- apply( KbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(T, d$K)
KbyT<- sapply(1:25, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_Y * colMeans(post$Y) + post$bK_dK * d$dK + colMeans(post$bK_T)[i] ))
PIs<- apply( KbyT , 2 , PI )
means<- apply( KbyT , 2 , mean )
points(1:25, means, col = 2, lwd = 4)
points(1:25, PIs[1, ], col = 2, lty = 2 )
points(1:25, PIs[2, ], col = 2, lty = 2 )
for(i in 1:25) lines(c(i,i), c(PIs[1, i], PIs[2, i]), col = 2)


# Model S

plot(d$dS, d$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + post$bS_dS * i + post$bS_Y * colMeans(post$Y) + colMeans(post$bS_T)[d$T]))
means<- apply( SbydS , 2 , mean )
PIs<- apply( SbydS , 2 , PI )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$S)
Yseq<-seq(-16, -2.4, len = 20)
SbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aS + post$bS_dS * d$dS + post$bS_Y * i + colMeans(post$bS_T)[d$T]))
means<- apply( SbyY , 2 , mean )
PIs<- apply( SbyY , 2 , PI )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(T, d$S)
SbyT<- sapply(1:25, 
    function (i) inv_logit(post$aS + post$bS_dS * d$dS + post$bS_Y * colMeans(post$Y) + colMeans(post$bS_T)[i]))
PIs<- apply( SbyT , 2 , PI )
means<- apply( SbyT , 2 , mean )
points(1:25, means, col = 2, lwd = 4)
points(1:25, PIs[1, ], col = 2, lty = 2)
points(1:25, PIs[2, ], col = 2, lty = 2)
for(i in 1:25) lines(c(i,i), c(PIs[1, i], PIs[2, i]), col = 2)

```


Looks great, let me compare all those models.


```{r comp2, fig.show="hold", out.width="50%"}
# Compare models
model_comparison <- compare(m1, m3)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison)

model_comparison <- compare(m2, m3)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison)
```

This looks Model 1 and 3 are almost same, but model 3 is a bit better. 

# Future Suggestions 
- Removing causal relationship in each model (for example, no causal relationship from Y to K) may be examined with actual data. 
- The Model 3 still has a problem of bK_T and bS_T as they do not fit the data. I am trying to understand the reason and fix it. 
```{r forgetit}
rm(list=ls())
```