# Model 1 

- Package Installation
```{r packages, error = FALSE, warning = FALSE, message = FALSE, results = "hide"}
library("rethinking")
library(dplyr)
library(dagitty)
par(mfrow = c(2, 2))
```

## Causal effects of the DAG 
```{r DAG}
kdag<-dagitty("dag {
  S-> K <- T 
  T -> S
  T -> SL -> K
}")
adjustmentSets(kdag, exposure="SL", outcome = "K")# This shows the backdoor path and how to stratify

```


## Synthetic Data
- create a synthetic data with assumptions. See overview.pdf to check the data relationship. 
```{r data, fig.show="hold", out.width="50%"}

# Function of SL by Y 
myFunc <- function(x) {
case_when(
    -16.000 <= x & x <= -10   ~ (x + 16) * 30 / 6 - 80,
    -10.000 < x  & x <= -7.300 ~ (x + 10) * 60 / 3 - 50,
    -7.300 < x   & x <= -6    ~ 4,
    -6.000 < x   & x <= -3.000 ~ (x + 6) * (-5) / 3 + 4,
    -3.000 < x   & x <= -2.400 ~ (x + 3) * (-5) / 0.5 - 1,
    TRUE                      ~ 1 
  )
}
curve(myFunc(x), col = 2, xlim=c(-16.000, -2.400), ylim=c(-85, 10))


# Initialize Y 
Y<- runif(200, -16, -2.4)

# T and SL caused by Y
T<- NULL
SL<-NULL
for (i in 1:length(Y)) {
    T[i]<-rpois(1,  (Y[i]+16)/13.6 * 25)
    if(T[i] < 1) {T[i]<- 1}
    else if(T[i] > 25) {T[i]<- 25} # Make sure T is between 1 and 25
    SL[i]<- rnorm(1, myFunc(Y[i]), 5) 
}

SL <- (SL - min(SL)) / (max(SL) - min(SL)) # to estimate easily, make it between 0 and 1 

# Check the data
plot(T, SL)
plot(Y, T)
plot(Y, SL)

# initialize dS with random numbers
dS<-rbeta(length(T), 1.2,1) * 200 
# S is affected by Y and dS (omit T as T and S are correlated by Y)
S<-NULL
for(i in 1:length(T))S[i] <- rnorm(1, 0.3*Y[i] -dS[i]/50, 0.1)/25 + 0.4

# Check the data
range(S) # try to make less than 0.5
plot(density(S))
plot(dS, S)
points(dS[T==2], S[T==2], col = 2, lwd =4)
points(dS[T==13], S[T==13], col = 3, lwd =4)
points(dS[T==7], S[T==7], col = 4, lwd =4)
plot(T, S)
plot(Y, S)
plot(T, dS) # no correlation 
plot(SL, S)

# Initialize dK with random numbers
dK<-rbeta(length(T), 1.4,1)*400
# K is affected by SL, Y, S, dK
K<-NULL
for(i in 1:length(T))K[i] <- rnorm(1, SL[i] + 0.1* Y[i] + S[i]- dK[i]/250, 0.2)/9 + 0.4

# Check the data
range(K)# try to make less than 0.5
plot(density(K))
plot(dK, K)
plot(SL, K)
plot(Y, K)
plot(S, K)
plot(dS, K)

# Standardize data without SL, S, K (those are ratio)
d<-list(T=T, dS=standardize(dS), S=S, K=K, dK=standardize(dK))

# check the data
plot(density(d$K))
plot(d$dK, d$K)
plot(SL, d$K)
plot(d$T, d$K)
plot(d$S, d$K)
plot(d$dS, d$S)
plot(Y, d$S)
plot(d$dS, d$K) # no correlation

```


## Prior Prediction 
- Also see overview.pdf to know statistical model and prior
- There are a few messages to warning my priors. 
- This would be due to the prior of Y, which is very weakly initialized. Since it only shows up during first iteration of warmup, we can safely ignore it. 
- If you really fix it, you can use normal distribution with large sigma and constraints between -16 and -2.4
```{r m1}
# Model 1 
m1<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_Y[T] * Y[i] + bK_dK * dK,

        # We do not know anything about this but think similar for all group. 
        # So use partial pooling to regularize the estimate. 
        # mean should start at 0 and has small variation.
        
        # In this case (m1), we do not include correlation. 
        # This is because multilevel models of S and SL includes the information of Y 
        # But we have to regularize. 
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_S <- bK_S_bar + sigma_bK_S * z_bK_S[T],
        vector [25]: bK_SL <- bK_SL_bar + sigma_bK_SL * z_bK_SL[T],
        vector [25]: bK_Y <- bK_Y_bar + sigma_bK_Y * z_bK_Y[T],
        c(aK_bar, bK_S_bar,  bK_SL_bar, bK_Y_bar) ~ normal(0,0.3),
        c(sigma_aK, sigma_bK_S, sigma_bK_SL, sigma_bK_Y) ~ exponential(6), # want low variation among groups
        z_aK[T] ~ normal(0,0.01),
        z_bK_S[T] ~ normal(0,0.01),
        z_bK_SL[T] ~ normal(0,0.01),
        z_bK_Y[T] ~ normal(0,0.01),

        # don't need regularization as dK is not affected by T
        # We are hoping/believing b_dK is negative, but this is not prior. 
        bK_dK ~ normal(0, 0.5),
        U ~ exponential(1),


    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_dS * dS + bS_Y[T] * Y[i],

        # I am hoping/believing b_dS is negative and b_Y_S is positive 
        # partial pooling for the same reason for a2 and b_Y_S
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],
        c(aS_bar, bS_Y_bar) ~ normal(0,0.1), 
        c(sigma_aS, sigmabS_Y) ~ exponential(6),
        z_aS[T] ~ normal(0,0.05),
        zbS_Y[T] ~ normal(0,0.05),

        # don't need regularization
        bS_dS ~ normal(0,0.1),
        sigma_S ~ exponential(1),


        # Model SL by Y 
        # The log function and prior was refered to the previous research of Sea Level Change over time. 
        # This function is abstruct so I will have to actually stan code to provide accurate estimates
        # ulam code in r does not allow us to hartd code the conditional function. 
        # vector[200]: SL ~ normal(muSL + aSL2, sigmaSL),
        vector[200]: SL ~ normal(mu_SL, sigma_SL),
        vector[200]: logit(mu_SL) <- aSL + Y[i] * bSL,
        aSL ~ normal(8, 0.01), # very low variation since this part is hard coded. 
        bSL ~ normal(0.8, 0.01), # very low variation since this part is hard coded. 

        sigma_SL ~ exponential(1), 

         # Model T, This should be hard coded based on the previous research
        T ~ poisson((Y + 16)/ ( -2.4 + 16) * 25),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1" # SL is between 0 and 1 
        #, T    = "lower=1,upper=25"  #this is unnecessary since T is provided in the dat 
    ),chains=4, cores=4, log_lik=TRUE
)


dashboard(m1)
precis(m1, depth=1)
```

# Posterior Check
```{r post}
post <- extract.samples(m1)
```

## Hidden parameters 
- The hidden parameters are estimated by the model constraints and distributions. 
- It is expected that the difference is small and has similar trend as the original data.
```{r hidden, fig.show="hold", out.width="50%"}
# Posterior Y changes 
plot(1:200, Y)
points(1:200, post$Y[1,], col = 2)
for(i in 1 : 200) lines(c(i, i), c(post$Y[1,i], Y[i]), col = 2)

plot(Y, d$K)
points(post$Y[1,], d$K, col = 2)
for(i in 1 : 200) lines(c(post$Y[1,i], Y[i]), c(d$K[i], d$K[i]), col = 2)

plot(SL, d$K)
points(post$SL[1,], d$K, col = 2)
for(i in 1 : 200) lines(c(post$SL[1,i], SL[i]), c(d$K[i], d$K[i]), col = 2)
points(post$SL[2,], d$K, col = 3)
for(i in 1 : 200) lines(c(post$SL[2,i], SL[i]), c(d$K[i], d$K[i]), col = 3)

plot(Y, SL)
points(post$Y[1,], post$SL[1,], col=2)
for(i in 1 : 200) lines(c(Y[i], post$Y[1,i]), c(SL[i], post$SL[1,i]), col = 2)


plot(Y, d$T)
points(post$Y[1,], d$T ,col = 2)
for (i in 1:200) lines(c(Y[i], post$Y[1,i]), c(d$T[i],d$T[i]), col = 2)

```


=> this shows that the difference is not very large and hidden parameters have similar trend as the original (synthetized) data. 

## Posterior Simulation 

### Model K 
```{r postpredK, fig.show="hold", out.width="50%"}
plot(d$S, d$K)
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$aK + post$bK_S * i + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(SL, d$K)
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, # Tricky but t and t to calculate estimates by group, though dK does not have group so do not need 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + post$bK_SL * i + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d$dK, d$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * i )) 
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(Y, d$K)
Yseq<-seq(-16, -2.4, len = 20)
KbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + post$bK_Y * i + rep(post$bK_dK,25) * d$K )) 
PIs<- apply( KbyY , 2 , PI )
means<- apply( KbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
```

### Model S by dS and Y 
- By the way, there is no post-S exists as it seems absolutely same as post-SL
- would be because it is one of the predictors that i provided data 

```{r postpredS, fig.show="hold", out.width="50%"}
post$SL[1,] == post$S[1,] 

plot(d$dS, d$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * i + post$bS_Y * colMeans(post$Y))) # Used mean of each simulated data
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$S)
Yseq<-seq(-16, -2.4, len = 20)
SbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * d$dS + post$bS_Y * i))
means<- apply( SbyY , 2 , mean )
PIs<- apply( SbyY , 2 , PI )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)
```

- Functions catch the trend very well!

### Model SL by Y 

```{r postpredSL, fig.show="hold", out.width="50%"}
plot(Y, SL)
points(post$Y[1,], post$SL[1,], col = 2)
Yseq<-seq(-16, -2.4, len = 20)
SLbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aSL + post$bSL * i))
PIs<- apply( SLbyY , 2 , PI )
means<- apply( SLbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

```


- The model 1 looks good. All posterior prediction plots shows inclusion of data points inside the model scope. 

- but also I am wondering if I use correlation between S and SL instead of modeling Y in K model. 
This is the m2


# Model 2 (m2)
- see priors in overview.pdf, but this include correlation between SL and S
- warnings are annoying but relatively few. 
```{r m2}
# Model 2
m2<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_Y[T] * Y[i] + bK_dK * dK,

        # We do not know anything about this but think similar for all group. 
        # So use partial pooling to regularize the estimate. 
        # mean should start at 0 and has small variation.
        
        # In this case (m2), we do include correlation between S and SL due to Y. 
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_Y <- bK_Y_bar + sigma_bK_Y * z_bK_Y[T],
        c(aK_bar, bK_Y_bar)~normal(0,0.3),
        c(sigma_aK, sigma_bK_Y) ~ exponential(6),
        z_aK[T]~normal(0,0.01),
        z_bK_Y[T] ~ normal(0,0.01),

        transpars>vector[25]:bK_S <<- v[,1],
        transpars>vector[25]:bK_SL <<- v[,2],

        matrix[25, 2]:v ~ multi_normal(c(bK_S_bar, bK_SL_bar), RhoSSL, sigmaSSL),
        c(bK_S_bar, bK_SL_bar) ~ normal(0, 0.3),
        corr_matrix[2]:RhoSSL ~ lkj_corr(1),# this is not too strong but good correlation
        vector[2]: sigmaSSL ~ exponential(2), 


        # don't need regularization
        # We are hoping/believing b_dK is negative, but this is not prior. 
        bK_dK ~ normal(0, 0.5),
        U ~ exponential(1),


    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_dS * dS,

        # I am hoping/believing b_dS is negative and b_Y_S is positive 
        # partial pooling for the same reason for a2 and b_Y_S
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],

        c(aS_bar, bS_Y_bar) ~ normal(0,0.3), 
        c(sigma_aS, sigmabS_Y) ~ exponential(6),
        z_aS[T] ~ normal(0,0.01),
        zbS_Y[T] ~ normal(0,0.01),

        # don't need regularization
        bS_dS ~ normal(0,0.1),
        sigma_S ~ exponential(1),


        # Model SL by Y 
        # The log function and prior was refered to the previous research of Sea Level Change over time. 
        # This function is abstruct so I will have to actually stan code to provide accurate estimates
        # ulam code in r does not allow us to hartd code the conditional function. 
        # vector[200]: SL ~ normal(muSL + aSL2, sigmaSL),
        vector[200]: SL ~ normal(mu_SL, sigma_SL),
        vector[200]: logit(mu_SL) <- aSL + Y[i] * bSL,
        aSL ~ normal(8, 0.01), # very low variation since this part is hard coded. 
        bSL ~ normal(0.8, 0.01), # very low variation since this part is hard coded. 

        sigma_SL ~ exponential(1), 

         # Model T 
        T ~ poisson((Y + 16)/ ( -2.4 + 16) * 25),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1" # SL is between 0 and 1 
    ),chains=4, cores=4, log_lik=TRUE 
)


dashboard(m2)
precis(m2, depth=1)
```

-> Looks worse sampling efficiency. 

# Posterior Check
```{r post2}
post <- extract.samples(m2)
```
- I do not check the Hidden parameters since I have never touched it since m1

## Correlation?
Rho value shows strong negative correlation between them. 
```{r post2cor, fig.show="hold", out.width="50%"}
R <- rlkjcorr( 1e4 , K=2 , eta=1 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1))
dens(post$Rho[,1,2], add=TRUE, lty=2) # low correlation
```
## Posterior Predictions

```{r post2pred, fig.show="hold", out.width="50%"}
plot(d$S, d$K)
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$aK + post$bK_S * i + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(SL, d$K)
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, # Tricky but t and t to calculate estimates by group, though dK does not have group so do not need 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + post$bK_SL * i + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * d$dK )) 
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(d$dK, d$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + t(t(post$bK_Y) * colMeans(post$Y)) + rep(post$bK_dK,25) * i )) 
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$K)
Yseq<-seq(-16, -2.4, len = 20)
KbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + post$bK_Y * i + rep(post$bK_dK,25) * d$K )) 
PIs<- apply( KbyY , 2 , PI )
means<- apply( KbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

# Model S
plot(d$dS, d$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * i )) # Used mean of each simulated data
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


```

K estimation by SL is a bit narrower than it supposed to be but other than that, everything looks great. 
Let me compare those two models 

```{r comp, fig.show="hold", out.width="50%"}
# Compare models
model_comparison <- compare(m1, m2)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison)
```

### Key Columns in the compare Output:
- WAIC / PSIS: Lower values indicate better out-of-sample predictive accuracy.
- dWAIC / dPSIS: The difference in information criteria between the current model and the top-ranked (best) model.
- dSE: If the standard error is large relative to the difference, the two models have statistically indistinguishable predictive capacities.
- weight: The Akaike weight, which represents the relative probability or evidence that a given model will make the best predictions among the set of compared models

Therefore, the m1 is better than the other one. 

# Model 3

```{r m3}
m3<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK + bK_S * S + bK_SL * SL[i] + bK_Y * Y[i] + bK_dK * dK + bK_T[T],

        # We do not know anything about this but think similar for all group. 
        # So use partial pooling to regularize the estimate. 
        # mean should start at 0 and has small variation.
        
        c(aK, bK_S,  bK_SL, bK_Y, bK_dK) ~ normal(0,0.3),
        vector[25]: bK_T <- bK_T_bar + sigma_bK_T * z_bK_T[T],
        bK_T_bar ~ normal(0, 0.3), # regularization
        sigma_bK_T ~ exponential(6), 
        z_bK_T[T] ~ normal(0, 1),

        U ~ exponential(1),


    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS + bS_dS * dS + bS_Y * Y[i] + bS_T[T],

        # I am hoping/believing b_dS is negative and b_Y_S is positive 
        # partial pooling for the same reason for a2 and b_Y_S
        c(aS, bS_Y, bS_dS) ~ normal(0,0.1), 
        vector[25]: bS_T <- bS_T_bar + sigma_bS_T * z_bS_T[T],
        bS_T_bar ~ normal(0, 0.001), # regularization
        sigma_bS_T ~ exponential(10), 
        z_bS_T[T] ~ normal(0, 0.1),

        sigma_S ~ exponential(1),

        # Model SL by Y 
        # The log function and prior was refered to the previous research of Sea Level Change over time. 
        # This function is abstruct so I will have to actually stan code to provide accurate estimates
        # ulam code in r does not allow us to hartd code the conditional function. 
        # vector[200]: SL ~ normal(muSL + aSL2, sigmaSL),
        vector[200]: SL ~ normal(mu_SL, sigma_SL),
        vector[200]: logit(mu_SL) <- aSL + Y[i] * bSL,
        aSL ~ normal(8, 0.01), # very low variation since this part is hard coded. 
        bSL ~ normal(0.8, 0.01), # very low variation since this part is hard coded. 

        sigma_SL ~ exponential(1), 

         # Model T 
        T ~ poisson((Y + 16)/ ( -2.4 + 16) * 25),

        # Model T by Y 
        vector[200]:Y ~ uniform( -16 , -2.4 )

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1" # SL is between 0 and 1 
    ),chains=4, cores=4, log_lik=TRUE
)

dashboard(m3)
precis(m3, depth=1)

```

-> Looks worse sampling efficiency. 

# Posterior Check
```{r post3}
post <- extract.samples(m3)
```
- Again, I do not check the Hidden parameters since I have never touched it since m1
```{r post3check, fig.show="hold", out.width="50%"}

# Model K

plot(d$S, d$K)
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$aK + post$bK_S * i + post$bK_SL * colMeans(post$SL) + post$bK_Y * colMeans(post$Y)  + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] ))
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(SL, d$K)
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq,  # This is tricky but use bK_T for T value and others are not stratified so randomly assigned
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * i + post$bK_Y * colMeans(post$Y)  + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] ))
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d$dK, d$K)
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_Y * colMeans(post$Y)  + post$bK_dK * i + colMeans(post$bK_T)[d$T] ))
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$K)
Yseq<-seq(-16, -2.4, len = 20)
KbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_Y * i + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] ))
PIs<- apply( KbyY , 2 , PI )
means<- apply( KbyY , 2 , mean )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(T, d$K)
KbyT<- sapply(1:25, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_Y * colMeans(post$Y) + post$bK_dK * d$dK + colMeans(post$bK_T)[i] ))
PIs<- apply( KbyT , 2 , PI )
means<- apply( KbyT , 2 , mean )
points(1:25, means, col = 2, lwd = 4)
points(1:25, PIs[1, ], col = 2, lty = 2 )
points(1:25, PIs[2, ], col = 2, lty = 2 )
for(i in 1:25) lines(c(i,i), c(PIs[1, i], PIs[2, i]), col = 2)


# Model S

plot(d$dS, d$S)
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + post$bS_dS * i + post$bS_Y * colMeans(post$Y) + colMeans(post$bS_T)[d$T]))
means<- apply( SbydS , 2 , mean )
PIs<- apply( SbydS , 2 , PI )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(Y, d$S)
Yseq<-seq(-16, -2.4, len = 20)
SbyY<- sapply(Yseq, 
    function (i) inv_logit(post$aS + post$bS_dS * d$dS + post$bS_Y * i + colMeans(post$bS_T)[d$T]))
means<- apply( SbyY , 2 , mean )
PIs<- apply( SbyY , 2 , PI )
lines(Yseq, means, col = 2, lwd = 3)
lines(Yseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Yseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(T, d$S)
SbyT<- sapply(1:25, 
    function (i) inv_logit(post$aS + post$bS_dS * d$dS + post$bS_Y * colMeans(post$Y) + colMeans(post$bS_T)[i]))
PIs<- apply( SbyT , 2 , PI )
means<- apply( SbyT , 2 , mean )
points(1:25, means, col = 2, lwd = 4)
points(1:25, PIs[1, ], col = 2, lty = 2)
points(1:25, PIs[2, ], col = 2, lty = 2)
for(i in 1:25) lines(c(i,i), c(PIs[1, i], PIs[2, i]), col = 2)

```


Looks great, let me compare all those models.


```{r comp2, fig.show="hold", out.width="50%"}
# Compare models
model_comparison <- compare(m1, m3)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison)

model_comparison <- compare(m2, m3)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison)
```

This looks Model 1 and 3 are almost same, but model 3 is a bit better. 

# Future Suggestions 
- Removing causal relationship in each model (for example, no causal relationship from Y to K) may be examined with actual data. 
- The Model 3 still has a problem of bK_T and bS_T as they do not fit the data. I am trying to understand the reason and fix it. 
```{r forgetit}
rm(list=ls())
```