---
title: "Models Rmd"
output:
  html_document: default
  pdf_document: default
---
# Models

- Package Installation
```{r packages, error = FALSE, warning = FALSE, message = FALSE, results = "hide"}
library("rethinking")
library(dplyr)
library(dagitty)
```

## Causal effects of the DAG 
```{r DAG}
kdag<-dagitty("dag {
  S-> K <- T 
  T -> S
  T -> SL -> K
}")
adjustmentSets(kdag, exposure="SL", outcome = "K")# This shows the backdoor path and how to stratify

```


## Synthetic Data
- create a synthetic data with assumptions. See overview.pdf to check the data relationship. 
```{r data, fig.show="hold", out.width="50%"}


# Function of SL by Y 
# This is arbitrary made by me, immitating the research by Susumu Tanabe and Rei Nakashima in 2026
# Y is between 0 and 1, which represents 16,000 years ago to 2,400 years ago. 
myFunc <- function(x) {
case_when(
    0 <= x & x <= (-9+ 16)/(13.6)                          ~ -(x * (-13.6)) * 60 / 7 - 80,
    (-9.000+ 16)/(13.6) < x  & x <= (-7.500 + 16)/(13.6)   ~ -(x * (-13.6) + 7) * 46 / 3 - 20,
    (-7.500+ 16)/(13.6) < x   & x <= (-4.3+ 16)/(13.6)        ~ 3,
    (-4.3 + 16)/(13.6) < x   & x <= (-3.000+ 16)/(13.6)    ~ -(x * (-13.6) + 11.7) * (-60) / 13 + 3,
    (-3.000+ 16)/(13.6) < x   & x <= (-2.400+ 16)/(13.6)    ~ -(x * (-13.6) + 13) * 5 / 3 - 1,
    TRUE                                                    ~ 1 
  )
}
curve(myFunc(x), col = 2, main = "Relative Sea Level and Year") 

# Initialize Y 
Y<- runif(200, 0, 1)

# T and SL caused by Y
T<- NULL
SL<-NULL
for (i in 1:length(Y)) {
    T[i]<-rpois(1,  Y[i]* 25)
    if(T[i] < 1) {T[i]<- 1}
    else if(T[i] > 25) {T[i]<- 25} # Make sure T is between 1 and 25
    SL[i]<- rnorm(1, myFunc(Y[i]), 5) 
}

SL <- (SL - min(SL)) / (max(SL) - min(SL)) # to estimate easily, make it between 0 and 1 

# Check the data
plot(T, SL, main = "Synthetic Data Plot(T and SL)")
plot(Y, T, main = "Synthetic Data Plot(Y and T)")
plot(Y, SL, main = "Synthetic Data Plot(Y and SL)")

# initialize dS with random numbers
dS<-rbeta(length(T), 1.2,1) * 200 
# S is affected by Y and dS (omit T as T and S are correlated by Y)
S<-NULL
for(i in 1:length(T))S[i] <- rnorm(1, 5*Y[i] -dS[i]/50, 0.1)/25 + 0.3

# Check the data
range(S) # try to make less than 0.5
plot(density(S), main = "Density of Synthetic Data of S")

plot(dS, S, main = "Synthetic Data Plot(dS and S) \n with different groups")
legend(x = 0, y = 0.18, legend="red: T = 2 blue: T = 7 green: T = 13")
points(dS[T==2], S[T==2], col = 2, lwd =4)
points(dS[T==13], S[T==13], col = 3, lwd =4)
points(dS[T==7], S[T==7], col = 4, lwd =4)
plot(T, S, main = "Synthetic Data Plot(T and S)")
plot(Y, S, main = "Synthetic Data Plot(Y and S)")
plot(T, dS, main = "Synthetic Data Plot(T and dS)\n No correlation expected.") # no correlation 
plot(SL, S, main = "Synthetic Data Plot(SL and S)\n auto-correlation")

# Initialize dK with random numbers
dK<-rbeta(length(T), 1.4,1)*400
# K is affected by SL, Y, S, dK
K<-NULL
for(i in 1:length(T))K[i] <- rnorm(1, 1 * SL[i] + 1 * Y[i] + 1 * S[i]- dK[i]/250, 0.2)/10 + 0.2

# Check the data
range(K)# try to make less than 0.5
plot(density(K), main = "Density of Synthetic Data of K")
plot(dK, K, main = "Synthetic Data Plot(dK and K)")
plot(SL, K, main = "Synthetic Data Plot(SL and K)")
plot(Y, K, main = "Synthetic Data Plot(Y and K)")
plot(S, K, main = "Synthetic Data Plot(S and K)")
plot(dS, K, main = "Synthetic Data Plot(dS and K)\n No correlation expected.") # no correlation

# Standardize data without SL, S, K (those are ratio)
d<-list(T=T, dS=standardize(dS), S=S, K=K, dK=standardize(dK))

# check the data
plot(density(d$K), main = "Density of Synthetic Data of d$K")
plot(d$dK, d$K, main = "Synthetic Data Plot(d$dK and d$K)")
plot(SL, d$K, main = "Synthetic Data Plot(SL and d$K)")
plot(Y, d$K, main = "Synthetic Data Plot(Y and d$K)")
plot(d$T, d$K, main = "Synthetic Data Plot(d$T and d$K)")
plot(d$S, d$K, main = "Synthetic Data Plot(d$S and d$K)")
plot(d$dS, d$S, main = "Synthetic Data Plot(d$dS and d$S)")
plot(Y, d$S, main = "Synthetic Data Plot(Y and d$S)")
plot(d$dS, d$K, main = "Synthetic Data Plot(d$dS and d$K)\n No correlation expected.") # no correlation

```

## Prior Prediction 
- Now I want to create a prior and statistical model. 
- As in the overview, the Y and Sl are hidden parameter so how to initialize is important
- This time, to make it simple, I use logit function to immitate the $myFunction$ 
```{r prior, fig.show="hold", out.width="50%"}
plot(Y, SL, main = "Synthetic Data Plot(Y and SL)")
# Simulation 
for(i in 1:10)curve(inv_logit(rnorm(1, 6, 0.2)*x + rnorm(1, -2, 0.2)) - rnorm(1, 0.04, 0.02), add =TRUE, col = 2)
curve(inv_logit(6*x - 2) - 0.04, add =TRUE, lwd = 3)
legend(0.5, 0.2,legend="Black line is my estimation, \nred is simulation")

```

## Model 1
- Also see overview.pdf to know statistical model and prior
- This model stratifies all parameters by T. 
- There are a few messages to warning my priors. Since it only shows up during first iteration of warmup, we can safely ignore it. 
- I just omit it for m1.2 and 1.3.

```{r m1}
# Model 1 
m1<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_Y[T] * Y[i] + bK_dK * dK,

        # We do not know anything about this but think similar for all group. 
        # So use partial pooling to regularize the estimate. 
        # mean should start at 0 and has small variation.
    
        # In this case (m1), we do not include correlation. 
        # This is because multilevel models of S and SL includes the information of Y 
        # But we have to regularize. 
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_S <- bK_S_bar + sigma_bK_S * z_bK_S[T],
        vector [25]: bK_SL <- bK_SL_bar + sigma_bK_SL * z_bK_SL[T],
        vector [25]: bK_Y <- bK_Y_bar + sigma_bK_Y * z_bK_Y[T],
        c(aK_bar, bK_S_bar,  bK_SL_bar, bK_Y_bar) ~ normal(0,0.3),
        c(sigma_aK, sigma_bK_S, sigma_bK_SL, sigma_bK_Y) ~ exponential(6), # want low variation among groups
        z_aK[T] ~ normal(0,0.1),
        z_bK_S[T] ~ normal(0,0.1),
        z_bK_SL[T] ~ normal(0,0.1),
        z_bK_Y[T] ~ normal(0,0.1),

        # don't need regularization as dK is not affected by T
        # We are hoping/believing b_dK is negative, but this is not prior. 
        bK_dK ~ normal(0, 0.3),
         U ~ exponential(1),

     # Model S 
         S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_Y[T] * Y[i] + bS_dS * dS,

         # I am hoping/believing b_dS is negative and b_Y_S is positive 
         # partial pooling for the same reason for a2 and b_Y_S
         vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
         vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],
         c(aS_bar, bS_Y_bar) ~ normal(0,0.3), 
         c(sigma_aS, sigmabS_Y) ~ exponential(6),
         z_aS[T] ~ normal(0,0.1),
         zbS_Y[T] ~ normal(0,0.1),

        # don't need regularization
         bS_dS ~ normal(0,0.3),
         sigma_S ~ exponential(1),

    # Model SL by Y
        # The log function and prior was refered to the previous research of Sea Level Change over time. 
        # This function is abstruct so I will have to actually stan code to provide accurate estimates
        # ulam code in r does not allow us to hartd code the conditional function. 
        # vector[200]: SL ~ normal(muSL + aSL2, sigmaSL),
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

    # Model T, This should be hard coded based on the previous research
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal(0.5, 0.3)

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1", # SL is between 0 and 1 
        Y    = "lower=0,upper=1" # Y is between 0 and 1 
        #, T    = "lower=1,upper=25"  #this is unnecessary since T is provided in the dat 

    
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE
)
dashboard(m1)
```

- This has so much divergent and less effective sampling. 
- This would be due to the large number of parameters, including hidden estimated ones (SL and Y). I will reduce parameters to reliably interpret the result. 
    - First one would be Y as S and SL are more important for the model. 
        - by the way, SL and Y in model K change the efficiency in the highest rate. 


```{r m1.2, error = FALSE, warning = FALSE, message = FALSE}
# Model 1.2 
m1.2<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_dK * dK,
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_S <- bK_S_bar + sigma_bK_S * z_bK_S[T],
        vector [25]: bK_SL <- bK_SL_bar + sigma_bK_SL * z_bK_SL[T],
        c(aK_bar, bK_S_bar,  bK_SL_bar) ~ normal(0,0.3),
        c(sigma_aK, sigma_bK_S, sigma_bK_SL) ~ exponential(6), 
        z_aK[T] ~ normal(0,0.1),
        z_bK_S[T] ~ normal(0,0.1),
        z_bK_SL[T] ~ normal(0,0.1),
        bK_dK ~ normal(0, 0.3),
        U ~ exponential(1),

    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_Y[T] * Y[i] + bS_dS * dS ,
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],
        c(aS_bar, bS_Y_bar) ~ normal(0,0.3), 
        c(sigma_aS, sigmabS_Y) ~ exponential(6),
        z_aS[T] ~ normal(0,0.1),
        zbS_Y[T] ~ normal(0,0.1),
        bS_dS ~ normal(0,0.3),
        sigma_S ~ exponential(1),

        # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

        # Model T, This should be hard coded based on the previous research
        T ~ poisson(Y * 25),

        # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1" 
        , Y    = "lower=0,upper=1"  
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE
)
dashboard(m1.2)
```


This is still bad model as the eff is very small number 

```{r m1.3, error = FALSE, warning = FALSE, message = FALSE}
# Model 1.3 
m1.3<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i]  + bK_dK * dK,

        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_S <- bK_S_bar + sigma_bK_S * z_bK_S[T],
        vector [25]: bK_SL <- bK_SL_bar + sigma_bK_SL * z_bK_SL[T],
        c(aK_bar, bK_S_bar,  bK_SL_bar) ~ normal(0,0.3),
        c(sigma_aK, sigma_bK_S, sigma_bK_SL) ~ exponential(6), 
        z_aK[T] ~ normal(0,0.1),
        z_bK_S[T] ~ normal(0,0.1),
        z_bK_SL[T] ~ normal(0,0.1),
        bK_dK ~ normal(0, 0.3),
        U ~ exponential(1),

    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_dS * dS,
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        aS_bar ~ normal(0,0.3), 
        sigma_aS ~ exponential(6),
        z_aS[T] ~ normal(0,0.1),
        bS_dS ~ normal(0,0.3),
        sigma_S ~ exponential(1),

    # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),
    # Model T, This should be hard coded based on the previous research
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)

    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1", 
        Y    = "lower=0,upper=1" 
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE
)
dashboard(m1.3)
precis(m1.3, depth=1)
```

- This is very high eff and no divergent. This is great. This model is reliably sampled to be used for analysis!

### Posterior Check
```{r post}
post <- extract.samples(m1.3)
```

#### Hidden parameters 
- The hidden parameters are estimated by the model constraints and distributions. 
- It is expected that the difference is small and has similar trend as the original data.
```{r hidden, fig.show="hold", out.width="50%"}
# Posterior Y changes 
plot(1:200, Y, main = "Difference between original (black) and simulated Y (red)")
points(1:200, post$Y[1,], col = 2)
for(i in 1 : 200) lines(c(i, i), c(post$Y[1,i], Y[i]), col = 2)

plot(1:200, SL, main = "Difference between original (black) and simulated SL(red)")
points(1:200, post$SL[1,], col = 2)
for(i in 1 : 200) lines(c(i, i), c(post$SL[1,i], SL[i]), col = 2)

# check Y and SL with K 
plot(Y, d$K, main = "Difference between original (black) and simulated Y (red) on d$K")
points(post$Y[1,], d$K, col = 2)
for(i in 1 : 200) lines(c(post$Y[1,i], Y[i]), c(d$K[i], d$K[i]), col = 2)

plot(SL, d$K, main = "Difference between original (black) and simulated SL (red and green) on d$K")
points(post$SL[1,], d$K, col = 2)
for(i in 1 : 200) lines(c(post$SL[1,i], SL[i]), c(d$K[i], d$K[i]), col = 2)
points(post$SL[2,], d$K, col = 3)
for(i in 1 : 200) lines(c(post$SL[2,i], SL[i]), c(d$K[i], d$K[i]), col = 3)

# Y and T 
plot(Y, d$T, main = "Difference between original (black) and simulated Y (red) on d$T")
points(post$Y[1,], d$T ,col = 2)
for (i in 1:200) lines(c(Y[i], post$Y[1,i]), c(d$T[i],d$T[i]), col = 2)

# Y and SL both hidden parameters
plot(Y, SL, main = "Difference between original (black) and simulation (red) of Y and SL")
points(post$Y[1,], post$SL[1,], col=2)
for(i in 1 : 200) lines(c(Y[i], post$Y[1,i]), c(SL[i], post$SL[1,i]), col = 2)

```


=> this shows that the difference is not very large and hidden parameters have similar trend as the original (synthetized) data. 

#### Posterior Simulation 

##### Model K 
```{r postpredK, fig.show="hold", out.width="50%"}
plot(d$S, d$K, main = "Synthetic Data Plot(d$S and d$K)")
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, # Tricky but t and t to calculate estimates by group, though dK does not have group so do not need 
    function (i) inv_logit(post$aK + post$bK_S * i + t(t(post$bK_SL) * colMeans(post$SL)) + rep(post$bK_dK, 25) * d$dK )) 
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(SL, d$K, main = "Synthetic Data Plot(SL and d$K)")
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + post$bK_SL * i + rep(post$bK_dK, 25) * d$dK )) 
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d$dK, d$K, main = "Synthetic Data Plot(d$dK and d$K)")
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + rep(post$bK_dK, 25) * i )) 
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

```

- Looks nice! 

##### Model S
- Do the same thing for S 
- By the way, there is no post-S exists as it seems absolutely same as post-SL
- would be because it is one of the predictors that i provided in the data.  

```{r postpredS, fig.show="hold", out.width="50%"}
post$SL[1,] == post$S[1,] 

plot(d$dS, d$S, main = "Synthetic Data Plot(d$dS and d$S)")
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * i )) # Used mean of each simulated data
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

```

- Functions catch the trend very well!


- The model 1.3 looks good. All posterior prediction plots shows inclusion of data points inside the model scope. 

- but also I am wondering if I use correlation between S and SL instead of modeling Y in K model. 
This is the m2

```{r cor, fig.show="hold", out.width="50%"}
plot(SL, S, main = "Synthetic Data Plot(SL and S)\n auto-correlation")
```
This is because there is a correlation!!

## Model 2 (m2)
- see priors in overview.pdf, but this include correlation between SL and S
- This is based on the Ch 14.3 in Statistical Rethinking as SL and S are not centred
- warnings are annoying but relatively few so I just omit it for m2.2 and m.3. 
```{r m2}
# Model 2
m2<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_Y[T] * Y[i] + bK_dK * dK,

        # We do not know anything about each parameter but think similar for all group. 
        # So use partial pooling to regularize estimates. 
        # mean should start at 0 and has small variation.
        
        # In this case (m2), we do include correlation between S and SL due to Y. 
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        vector [25]: bK_Y <- bK_Y_bar + sigma_bK_Y * z_bK_Y[T],
        c(aK_bar, bK_Y_bar)~normal(0,0.3),
        c(sigma_aK, sigma_bK_Y) ~ exponential(6),
        z_aK[T]~normal(0,0.1),
        z_bK_Y[T] ~ normal(0,0.1),

        transpars>vector[25]:bK_S <<- v[,1],
        transpars>vector[25]:bK_SL <<- v[,2],

        transpars>matrix[25, 2]:v <- compose_noncentered(sigmaSSL, RhoSSL, zSSL), 
        vector[2]: sigmaSSL ~ exponential(2),
        cholesky_factor_corr[2]:RhoSSL ~ lkj_corr_cholesky(1), # I do not expect very very strong correlation.
        matrix[2,25]: zSSL ~ normal(0 , 1), 

        # don't need regularization
        # We are hoping/believing b_dK is negative, but this is not prior. 
        bK_dK ~ normal(0, 0.3),
        U ~ exponential(1),


    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_Y[T] * Y[i] + bS_dS * dS ,

        # I am hoping/believing b_dS is negative and b_Y_S is positive 
        # partial pooling for the same reason for a2 and b_Y_S
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],
        c(aS_bar, bS_Y_bar) ~ normal(0,0.3), 
        c(sigma_aS, sigmabS_Y) ~ exponential(6),
        z_aS[T] ~ normal(0,0.1),
        zbS_Y[T] ~ normal(0,0.1),

        # don't need regularization
        bS_dS ~ normal(0,0.3),
        sigma_S ~ exponential(1),

    # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

    # Model T, This should be hard coded based on the previous research
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)
    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1", # SL is between 0 and 1 
        Y    = "lower=0,upper=1" # Y is between 0 and 1 
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE, max_treedepth = 13, control = list(
    adapt_delta = 0.99
  )
)

dashboard(m2)
```

-> Looks bad sampling efficiency. 

Therefore, I will tackle this problem with the same approach though I do not want to remove the correlation parameters. 

```{r m2.2, error = FALSE, warning = FALSE, message = FALSE}
# Model 2,2
m2.2<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_dK * dK,
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        aK_bar~normal(0,0.3),
        sigma_aK ~ exponential(6),
        z_aK[T]~normal(0,0.1),

        transpars>vector[25]:bK_S <<- v[,1],
        transpars>vector[25]:bK_SL <<- v[,2],

        transpars>matrix[25, 2]:v <- compose_noncentered(sigmaSSL, RhoSSL, zSSL), 
        vector[2]: sigmaSSL ~ exponential(2),
        cholesky_factor_corr[2]:RhoSSL ~ lkj_corr_cholesky(1),
        matrix[2,25]: zSSL ~ normal(0 , 1), 
        bK_dK ~ normal(0, 0.3),
        U ~ exponential(1),

    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_Y[T] * Y[i] + bS_dS * dS ,
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        vector[25]: bS_Y<-bS_Y_bar + sigmabS_Y * zbS_Y[T],
        c(aS_bar, bS_Y_bar) ~ normal(0,0.3), 
        c(sigma_aS, sigmabS_Y) ~ exponential(6),
        z_aS[T] ~ normal(0,0.1),
        zbS_Y[T] ~ normal(0,0.1),
        bS_dS ~ normal(0,0.3),
        sigma_S ~ exponential(1),

    # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

    # Model T
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)
    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1", 
        Y    = "lower=0,upper=1" 
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE, max_treedepth = 13, control = list(
    adapt_delta = 0.99
  )
)
dashboard(m2.2)
```

- This sampling efficiency is still not great. Let me remove one more thing so that it corresponds m1.3. 

```{r m2.3, error = FALSE, warning = FALSE, message = FALSE}
# Model 2.3
m2.3<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK[T] + bK_S[T] * S + bK_SL[T] * SL[i] + bK_dK * dK,
        vector[25]: aK <- aK_bar + sigma_aK * z_aK[T],
        aK_bar~normal(0,0.3),
        sigma_aK ~ exponential(6),
        z_aK[T]~normal(0,0.1),
        transpars>vector[25]:bK_S <<- v[,1],
        transpars>vector[25]:bK_SL <<- v[,2],
        transpars>matrix[25, 2]:v <- compose_noncentered(sigmaSSL, RhoSSL, zSSL), 
        vector[2]: sigmaSSL ~ exponential(2),
        cholesky_factor_corr[2]:RhoSSL ~ lkj_corr_cholesky(1),
        matrix[2,25]: zSSL ~ normal(0 , 1), 
        bK_dK ~ normal(0, 0.3),
        U ~ exponential(1),

    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS[T] + bS_dS * dS ,
        vector[25]: aS<-aS_bar + sigma_aS * z_aS[T],
        aS_bar ~ normal(0,0.3), 
        sigma_aS ~ exponential(6),
        z_aS[T] ~ normal(0,0.1),
        bS_dS ~ normal(0,0.1),
        sigma_S ~ exponential(1),

    # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

    # Model T
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)
    ), dat = d, constraints = list(
        SL    = "lower=0,upper=1", 
        Y    = "lower=0,upper=1" 
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE, max_treedepth = 13, control = list(
    adapt_delta = 0.99
  )
)
dashboard(m2.3)
precis(m2.3, depth=1)
```

- This model is good enough.

### Posterior Check
```{r post2}
post <- extract.samples(m2.3)
```

- I do not check the Hidden parameters since I have never touched it since m1

#### Correlation?
- Rho value shows a bit of correlation between S and SL. 
```{r post2cor, fig.show="hold", out.width="50%"}
R <- rlkjcorr( 1e4 , K=2 , eta=1 )# eta min lkj_corr()
plot(density( R[,1,2]), ylim = c(0, 1), main = "prior (black) and posterior (red) of Rho")
# This is complicated as cholesky correlation 
true_rho_samples=NULL
for (i in 1:2000) {
  L <- post$RhoSSL[i, , ]
  R <- L %*% t(L)
  true_rho_samples[i] <- R[1, 2]
}
dens(true_rho_samples, add = TRUE, lty = 2, col = 2)

```
#### Posterior Predictions

```{r post2pred, fig.show="hold", out.width="50%"}

plot(d$S, d$K, main = "Synthetic Data Plot(d$S and d$K)")
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, # Tricky but t and t to calculate estimates by group, though dK does not have group so do not need 
    function (i) inv_logit(post$aK + post$bK_S * i + t(t(post$bK_SL) * colMeans(post$SL)) + rep(post$bK_dK, 25) * d$dK )) 
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(SL, d$K, main = "Synthetic Data Plot(SL and d$K)")
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + post$bK_SL * i + rep(post$bK_dK, 25) * d$dK )) 
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d$dK, d$K, main = "Synthetic Data Plot(d$dK and d$K)")
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + t(t(post$bK_S) * d$S) + t(t(post$bK_SL) * colMeans(post$SL)) + rep(post$bK_dK, 25) * i )) 
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

# Model S
plot(d$dS, d$S, main = "Synthetic Data Plot(d$dS and d$S)")
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + rep(post$bS_dS, 25) * i )) # Used mean of each simulated data
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


```

Overall K estimation is fine, but fit less than m1.3. 

Let me compare those two models 

### Compare Models 
```{r comp, fig.show="hold", out.width="50%"}
# Compare models
model_comparison <- compare(m1.3, m2.3)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison, main = "Model Comparison between m1.3 and m2.3")
```

#### Key Columns in the compare Output:
- WAIC / PSIS: Lower values indicate better out-of-sample predictive accuracy.
- dWAIC / dPSIS: The difference in information criteria between the current model and the top-ranked (best) model.
- dSE: If the standard error is large relative to the difference, the two models have statistically indistinguishable predictive capacities.
- weight: The Akaike weight, which represents the relative probability or evidence that a given model will make the best predictions among the set of compared models

=> Therefore, the m1 is better than the other one. 


## Model 3
- Same thing! See Overview pdf for details. 
```{r m3}
m3<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK + bK_S * S + bK_SL * SL[i] + bK_Y * Y[i] + bK_dK * dK + bK_T[T],

        # Basically same as m1 but not regularization (no pooling). 
        # bK_T is regularized to maintain the efficiency. 

        c(aK, bK_S,  bK_SL, bK_Y, bK_dK) ~ normal(0,0.1), # This is different from the other one, so that bK_T works well; otherwise, it does not somehow.
        vector[25]: bK_T <- bK_T_bar + sigma_bK_T * z_bK_T[T],
        bK_T_bar ~ normal(0, 0.1), # regularization but not high sigma
        sigma_bK_T ~ exponential(1), # Needs to be 1 to maintain bK_T. 
        z_bK_T[T] ~ normal(0, 0.1),

        U ~ exponential(1),

    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS + bS_dS * dS + bS_Y * Y[i] + bS_T[T],

        # I am hoping/believing b_dS is negative, b_Y_S is positive, and bS_T is different for each one.  
        c(aS, bS_Y, bS_dS) ~ normal(0,0.1), 
        # bS_T is regularized. 
        vector[25]: bS_T <- bS_T_bar + sigma_bS_T * z_bS_T[T],
        bS_T_bar ~ normal(0, 0.1), 
        sigma_bS_T ~ exponential(1), 
        z_bS_T[T] ~ normal(0, 0.1),

        sigma_S ~ exponential(1),

    # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

    # Model T
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)
    ), dat = d, constraints = list(
        SL   = "lower=0,upper=1", # SL is between 0 and 1 
        Y    = "lower=0,upper=1" # Y is between 0 and 1 
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE, max_treedepth = 13, control = list(
    adapt_delta = 0.99
  )
)
dashboard(m3)

```

- without divergent but it has a few problem with efficiency. I will try to eliminate by removing parameters. 

```{r m3.2, error = FALSE, warning = FALSE, message = FALSE}
m3.2<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK + bK_S * S + bK_SL * SL[i] + bK_dK * dK + bK_T[T],

        c(aK, bK_S,  bK_SL, bK_Y, bK_dK) ~ normal(0,0.1),
        vector[25]: bK_T <- bK_T_bar + sigma_bK_T * z_bK_T[T],
        bK_T_bar ~ normal(0, 0.1), 
        sigma_bK_T ~ exponential(1), 
        z_bK_T[T] ~ normal(0, 0.1),

        U ~ exponential(1),

    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS + bS_dS * dS + bS_Y * Y[i] + bS_T[T],

        c(aS, bS_Y, bS_dS) ~ normal(0,0.1), 
        vector[25]: bS_T <- bS_T_bar + sigma_bS_T * z_bS_T[T],
        bS_T_bar ~ normal(0, 0.1), 
        sigma_bS_T ~ exponential(1), 
        z_bS_T[T] ~ normal(0, 0.1),

        sigma_S ~ exponential(1),

    # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

    # Model T
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)
    ), dat = d, constraints = list(
        SL   = "lower=0,upper=1", 
        Y    = "lower=0,upper=1"  
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE, max_treedepth = 13, control = list(
    adapt_delta = 0.99
  )
)

dashboard(m3.2)
```

- This is still not the best model with low efficiency. I will remove one more parameter. 
- I can remove bK_T or bS_T but this time, I want to make this similar to m1.3 and m2.3. 
```{r m3.3, error = FALSE, warning = FALSE, message = FALSE}
m3.3<- ulam(
    alist (
    # Main model K
        K ~ normal(mu_K, U), 
        logit(mu_K) <- aK + bK_S * S + bK_SL * SL[i] + bK_dK * dK + bK_T[T],

        c(aK, bK_S,  bK_SL, bK_dK) ~ normal(0,0.1),
        vector[25]: bK_T <- bK_T_bar + sigma_bK_T * z_bK_T[T],
        bK_T_bar ~ normal(0, 0.1), 
        sigma_bK_T ~ exponential(1), 
        z_bK_T[T] ~ normal(0, 0.1),

        U ~ exponential(1),

    # Model S 
        S ~ normal(mu_S, sigma_S),
        logit(mu_S) <- aS + bS_dS * dS + bS_T[T],

        c(aS, bS_dS) ~ normal(0,0.1), 
        vector[25]: bS_T <- bS_T_bar + sigma_bS_T * z_bS_T[T],
        bS_T_bar ~ normal(0, 0.1), 
        sigma_bS_T ~ exponential(1), 
        z_bS_T[T] ~ normal(0, 0.1),

        sigma_S ~ exponential(1),

    # Model SL by Y
        vector[200]: SL ~ normal(inv_logit(6 * Y - 2) - 0.04, 0.05),

    # Model T
        T ~ poisson(Y * 25),

    # Model T by Y 
        vector[200]:Y ~ normal( 0.5, 0.2)
    ), dat = d, constraints = list(
        SL   = "lower=0,upper=1",
        Y    = "lower=0,upper=1" 
    ), chains=4, cores=4, iter= 4000, warmup = 2000, log_lik=TRUE, max_treedepth = 13, control = list(
    adapt_delta = 0.99
  )
)

dashboard(m3.3)
precis(m3.3, depth = 1)

```

### Posterior Check
```{r post3}
post <- extract.samples(m3.3)
```
- Again, I do not check the Hidden parameters since I have never touched it since m1

```{r post3check, fig.show="hold", out.width="50%"}

plot(d$S, d$K, main = "Synthetic Data Plot(d$S and d$K)")
Sseq<-seq(0, 0.5, len = 20)
KbyS<- sapply(Sseq, 
    function (i) inv_logit(post$aK + post$bK_S * i + post$bK_SL * colMeans(post$SL) + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] ) )
means<- apply( KbyS , 2 , mean )
PIs<- apply( KbyS , 2 , PI )
lines(Sseq, means, col = 2, lwd = 3)
lines(Sseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(Sseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(SL, d$K, main = "Synthetic Data Plot(SL and d$K)")
points(post$SL[1,], d$K, col = 2) # include the estimated points, though we confirmed above that it is close to the orignial
SLseq<-seq(0, 1, len = 20)
KbySL<- sapply(SLseq, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * i + post$bK_dK * d$dK + colMeans(post$bK_T)[d$T] )) 
means<- apply( KbySL , 2 , mean )
PIs<- apply( KbySL , 2 , PI )
lines(SLseq, means, col = 2, lwd = 3)
lines(SLseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(SLseq, PIs[2, ], col = 2, lty = 2, lwd = 3)


plot(d$dK, d$K, main = "Synthetic Data Plot(d$dK and d$K)")
dKseq<-seq(-2.5, 2, len = 20)
KbydK<- sapply(dKseq, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_dK * i + colMeans(post$bK_T)[d$T] )) 
PIs<- apply( KbydK , 2 , PI )
means<- apply( KbydK , 2 , mean )
lines(dKseq, means, col = 2, lwd = 3)
lines(dKseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dKseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(d$T, d$K, main = "Synthetic Data Plot(d$T and d$K)")
KbyT<- sapply(1:25, 
    function (i) inv_logit(post$aK + post$bK_S * d$S + post$bK_SL * colMeans(post$SL) + post$bK_dK * d$dK + colMeans(post$bK_T)[i]))
PIs<- apply( KbyT , 2 , PI )
means<- apply( KbyT , 2 , mean )
points(1:25, means, col = 2, lwd = 4)
points(1:25, PIs[1, ], col = 2, lty = 2 )
points(1:25, PIs[2, ], col = 2, lty = 2 )
for(i in 1:25) lines(c(i,i), c(PIs[1, i], PIs[2, i]), col = 2)

# Model S
plot(d$dS, d$S, main = "Synthetic Data Plot(d$dS and d$S)")
dSseq<-seq(-2, 2, len = 20)
SbydS<- sapply(dSseq, 
    function (i) inv_logit(post$aS + post$bS_dS * i + colMeans(post$bS_T)[d$T] )) # Used mean of each simulated data
PIs<- apply( SbydS , 2 , PI )
means<- apply( SbydS , 2 , mean )
lines(dSseq, means, col = 2, lwd = 3)
lines(dSseq, PIs[1, ], col = 2, lty = 2, lwd = 3)
lines(dSseq, PIs[2, ], col = 2, lty = 2, lwd = 3)

plot(T, d$S, main = "Synthetic Data Plot(d$T and d$S)")
SbyT<- sapply(1:25, 
    function (i) inv_logit(post$aS + post$bS_dS * d$dS + colMeans(post$bS_T)[i]))
PIs<- apply( SbyT , 2 , PI )
means<- apply( SbyT , 2 , mean )
points(1:25, means, col = 2, lwd = 4)
points(1:25, PIs[1, ], col = 2, lty = 2)
points(1:25, PIs[2, ], col = 2, lty = 2)
for(i in 1:25) lines(c(i,i), c(PIs[1, i], PIs[2, i]), col = 2)

```


I have no idea why it does not model bK_T and bS_T very well... I will figure it out.
Other than that, looks great, let me compare all those models.

### Compare Models 

```{r comp2, fig.show="hold", out.width="50%"}
# Compare models
model_comparison <- compare(m1.3, m3.3)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison, main = "Model Comparison between m1.3 and m3.3")

model_comparison <- compare(m2.3, m3.3)
print(model_comparison)
# Plot WAIC/PSIS differences
plot(model_comparison, main = "Model Comparison between m2.3 and m3.3")
```

This looks Model 1 is a bit better in this case. 

## Future Suggestions 
- Removing causal relationship in each model (for example, no causal relationship from Y to K) may be examined with actual data. 
- The Model 3 still has a problem of bK_T and bS_T as they do not fit the data. I am trying to understand the reason and fix it. 
```{r forgetit}
rm(list=ls())
```